### 2020-01-21 SOCbudgetPaper/conclusion.Rmd

\conclusions

The conclusion goes here.
You can modify the section name with  `\conclusions[modified heading if necessary]`.