### 2020-04-02 ResultNotebooks/results_socbudget.tex
Update pdf and results

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={First SOC budget analysis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{First SOC budget analysis}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(moinput)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: magclass
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'magclass'
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:grid':
## 
##     getNames
\end{verbatim}

\begin{verbatim}
## Loading required package: madrat
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(mrvalidation)}
\KeywordTok{library}\NormalTok{(socbudget)}

\KeywordTok{setConfig}\NormalTok{(}\DataTypeTok{cachefolder =} \StringTok{"/home/kristine/WORK/R libraries/rd3mod\_inputdata/cache/rev4.18.cellpanda/"}\NormalTok{, }\DataTypeTok{forcecache =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Configuration update:
##   cachefolder: NA -> /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda
##   forcecache: FALSE -> TRUE
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")}
\KeywordTok{setConfig}\NormalTok{(}\DataTypeTok{sourcefolder =} \StringTok{"/home/kristine/WORK/R libraries/rd3mod\_inputdata/sources/"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Configuration update:
##   sourcefolder: NA -> /home/kristine/WORK/R libraries/rd3mod_inputdata/sources
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SoilCarbon <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"CarbonBudget"}\NormalTok{, }\DataTypeTok{aggregate =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("CarbonBudget", aggregate = FALSE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcCarbonBudget.Rda
## Exit calcOutput("CarbonBudget", aggregate = FALSE) in 10.32 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Landuse    <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"Landuse"}\NormalTok{,      }\DataTypeTok{aggregate=}\OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("Landuse", aggregate = FALSE)
\end{verbatim}

\begin{verbatim}
## Warning: More than one function found for type "Landuse" and prefix "calc". Use
## last occurrence (package "socbudget")
\end{verbatim}

\begin{verbatim}
## > WARNING: More than one function found for type " Landuse " and prefix " calc 
## > ". Use last occurrence (package " socbudget ")
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcLanduse.Rda
## Exit calcOutput("Landuse", aggregate = FALSE) in 1.87 seconds
\end{verbatim}

Calculate global carbon stocks

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SOCStock     <{-}}\StringTok{ }\KeywordTok{dimSums}\NormalTok{(}\KeywordTok{dimSums}\NormalTok{(SoilCarbon, }\DataTypeTok{dim=}\FloatTok{3.2}\NormalTok{)}\OperatorTok{*}\StringTok{ }\NormalTok{Landuse, }\DataTypeTok{dim=}\DecValTok{3}\NormalTok{)}
\NormalTok{glo\_SOCStock <{-}}\StringTok{ }\KeywordTok{dimSums}\NormalTok{(SOCStock, }\DataTypeTok{dim=}\DecValTok{1}\NormalTok{) }
\NormalTok{cell2iso     <{-}}\StringTok{ }\KeywordTok{toolGetMapping}\NormalTok{(}\StringTok{"CountryToCellMapping.csv"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"cell"}\NormalTok{)}
\NormalTok{iso\_SOCStock <{-}}\StringTok{ }\KeywordTok{toolCountryFill}\NormalTok{(}\KeywordTok{toolAggregate}\NormalTok{(SOCStock, }\DataTypeTok{rel=}\NormalTok{cell2iso, }\DataTypeTok{from=}\StringTok{"celliso"}\NormalTok{, }\DataTypeTok{to=}\StringTok{"iso"}\NormalTok{, }\DataTypeTok{dim=}\DecValTok{1}\NormalTok{), }\DataTypeTok{fill=}\DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## NOTE:  - toolCountryFill set missing values for IMPORTANT countries to  0 :
## NOTE:  --- Bahrain(BHR)  --- Hong Kong(HKG)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{iso2reg      <{-}}\StringTok{ }\KeywordTok{toolMappingFile}\NormalTok{(}\StringTok{"regional"}\NormalTok{,}\KeywordTok{getConfig}\NormalTok{(}\StringTok{"regionmapping"}\NormalTok{),}\DataTypeTok{readcsv =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{reg\_SOCStock <{-}}\StringTok{ }\KeywordTok{toolAggregate}\NormalTok{(iso\_SOCStock, }\DataTypeTok{rel=}\NormalTok{iso2reg, }\DataTypeTok{from=}\StringTok{"CountryCode"}\NormalTok{, }\DataTypeTok{to=}\StringTok{"RegionCode"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Load validation data:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  LPJ\_IPCC   <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"LPJ\_IPCC2006"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006",     aggregate = "REG+GLO", try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_LPJ_IPCC2006.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006",     aggregate = "REG+GLO", try = TRUE) in 0.34 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  LPJ\_rev21  <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"LPJmL\_rev21"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_LPJmL_rev21.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate = "REG+GLO",     try = TRUE) in 0.14 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  LPJ\_car    <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"LPJmLCarbon"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_LPJmLCarbon.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate = "REG+GLO",     try = TRUE) in 0.08 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  WISE       <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"WISE"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_WISE.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate = "REG+GLO",     try = TRUE) in 0.09 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  GSOC       <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"GSOC"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_GSOC.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate = "REG+GLO",     try = TRUE) in 0.09 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  SoilGrids  <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"SoilGrids"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_SoilGrids.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate = "REG+GLO",     try = TRUE) in 0.09 seconds
\end{verbatim}

Plot global stocks

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(glo\_SOCStock, }\DataTypeTok{as.integer=}\OtherTok{TRUE}\NormalTok{)), glo\_SOCStock, }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{, }\DataTypeTok{xlim=}\KeywordTok{c}\NormalTok{(}\DecValTok{1951}\NormalTok{,}\DecValTok{2017}\NormalTok{),  }\DataTypeTok{ylim=}\KeywordTok{c}\NormalTok{(}\DecValTok{400000}\NormalTok{,}\DecValTok{1300000}\NormalTok{))}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(LPJ\_IPCC, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), LPJ\_IPCC[}\StringTok{"GLO"}\NormalTok{,,}\DecValTok{1}\NormalTok{], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"green"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(LPJ\_car, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), LPJ\_car[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"darkgreen"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(LPJ\_rev21, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), LPJ\_rev21[}\StringTok{"GLO"}\NormalTok{,,}\DecValTok{2}\NormalTok{], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"lightgreen"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(WISE, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), WISE[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"brown"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(SoilGrids, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), SoilGrids[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"blue"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(GSOC, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), GSOC[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"pink"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{results_socbudget_files/figure-latex/plot glo stocks-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(glo\_SOCStock, }\DataTypeTok{as.integer=}\OtherTok{TRUE}\NormalTok{)), glo\_SOCStock, }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{results_socbudget_files/figure-latex/plot glo stocks-2.pdf}

Plot regional stocks

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(mip)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: quitte
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(quitte)}
\KeywordTok{library}\NormalTok{(ggplot2)}

\KeywordTok{getNames}\NormalTok{(reg\_SOCStock) <{-}}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\StringTok{"modeloutput.SOCbudget."}\NormalTok{, }\KeywordTok{getNames}\NormalTok{(WISE,}\DataTypeTok{dim=}\DecValTok{3}\NormalTok{))}
\KeywordTok{getSets}\NormalTok{(reg\_SOCStock)  <{-}}\StringTok{ }\KeywordTok{getSets}\NormalTok{(WISE) }

\NormalTok{hist <{-}}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{as.quitte}\NormalTok{(WISE[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(GSOC[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(SoilGrids[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(LPJ\_IPCC[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{][,,}\DecValTok{1}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(LPJ\_car[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(LPJ\_rev21[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{][,,}\DecValTok{2}\NormalTok{]))}

\KeywordTok{mipLineHistorical}\NormalTok{(reg\_SOCStock, hist, }\DataTypeTok{facet.ncol=}\DecValTok{4}\NormalTok{, }\DataTypeTok{scales=}\StringTok{"free\_y"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth]{results_socbudget_files/figure-latex/reg c stock-1}

\begin{Shaded}
\begin{Highlighting}[]
    \CommentTok{\#cell\_GSOC      <{-} readSource("GSOC",  convert="onlycorrect")}
    \CommentTok{\#cell\_WISE      <{-} readSource("WISE",  convert="onlycorrect")}
    \CommentTok{\#cell\_SoilGrids <{-} readSource("SoilGrids",  convert="onlycorrect")}

    \CommentTok{\#koeppen     <{-} readSource("LPJml\_rev21", subtype ="koeppen", convert="onlycorrect" )}
    \CommentTok{\#koepp2IPCC  <{-} toolGetMapping("mapping\_koeppen\_ipcc.csv", type = "sectoral")}
    \CommentTok{\#climateIPCC <{-} toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen\_geiger", to="ipcc\_reduced")}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cropland\_share <{-}}\StringTok{ }\KeywordTok{dimSums}\NormalTok{(SoilCarbon[,,}\StringTok{"crop"}\NormalTok{], }\DataTypeTok{dim=}\DecValTok{3}\NormalTok{)}\OperatorTok{/}\KeywordTok{dimSums}\NormalTok{(SoilCarbon[,,}\StringTok{"natveg"}\NormalTok{], }\DataTypeTok{dim=}\DecValTok{3}\NormalTok{)}

\NormalTok{luplot}\OperatorTok{::}\KeywordTok{plotmap}\NormalTok{(cropland\_share[,}\StringTok{"y2000"}\NormalTok{,])}
\end{Highlighting}
\end{Shaded}

\includegraphics{results_socbudget_files/figure-latex/plot density maps-1.pdf}

\end{document}