### 2020-05-03 ResultNotebooks/SOCflows_results.Rmd
methods update, add more result rmds

---
  title: "First SOC budget analysis"
output:
  html_document:
  df_print: paged
pdf_document: default
---

```{r load}
library(moinput)
library(mrvalidation)
library(socbudget)
library(magpiesets)

#input       <- "/home/kristine/WORK/R libraries/rd3mod_inputdata/output/rev1_690d3718e151be1b450b394c1064b1c5_validation.tgz"
#inputfolder <- gsub("[^/]*$","", input)
#untar(input, inputfolder)

setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")

```

```{r residues}

### Residues
  # originally from moinput
  ResidueDemand       <- dimSums(collapseNames(calcOutput("ResFieldBalancePast", cellular=TRUE, products = "kres", aggregate = FALSE)[,,"c"]), dim=1)
  ResidueRecyclingAg  <- ResidueDemand[,,"recycle"]
  ResidueRecyclingBg  <- dimSums(calcOutput("ResBiomass", cellular=TRUE, aggregate = FALSE)[,,"bg"][,,c("c")], dim=c(1,3))
  ResiduesMoinput     <- dimSums(ResidueRecyclingAg, dim=3) + ResidueRecyclingBg

  # Residues with cut of overflow
  ResiduesDensities   <- calcOutput("CarbonResidues", aggregate=FALSE)[,,"c"]
  Landuse             <- calcOutput("Landuse", aggregate=FALSE)
  ResiduesSOCbudget   <- dimSums(Landuse[,getYears(ResiduesDensities),"crop"]*ResiduesDensities, dim=c(1,3))

  print(mbind(ResiduesMoinput, ResiduesSOCbudget, ResiduesSOCbudget/ResiduesMoinput))

  # Corrected Residues shares for different plant parts  
  RatioPlantParts     <- mbind(ResidueRecyclingAg, setNames(ResidueRecyclingBg, "res_roots"))/(dimSums(ResidueRecyclingAg, dim=3)+ResidueRecyclingBg)

  ResiduesCorrected   <- collapseNames(RatioPlantParts * ResiduesSOCbudget)

  # total residues and residue demands

  ResidueBg     <- ResidueRecyclingBg
  ResidueAg     <- dimSums(calcOutput("ResBiomass", cellular=TRUE, aggregate = FALSE)[,,"ag"][,,c("c")], dim=c(1,3))
  ResidueTotal  <- ResidueAg + ResidueBg

```

Quite a lot of residues get lost, due to cutting extreme values in the later years.
Cutting algorithm, cuts everything bigger the 95%-percentile.
How we deal with that? 
* Should we do a yearly extrem value cutting?
* Should we investigate that deeper?
* Should we leave it like this?

```{r manure}

### Manure
  # originally from moinput
  ManureApplication   <- calcOutput("ManureRecyclingCroplandPast", products = "kli", cellular = TRUE, aggregate = FALSE)[,,"c"]
  ManureGrazing       <- calcOutput("Excretion", cellular = TRUE, attributes = "npkc", aggregate = FALSE)[,,"stubble_grazing"][,,"c"]
  ManureMoinput       <- collapseNames(dimSums(ManureApplication, dim=1) + dimSums(ManureGrazing, dim=1))

  # Manure with cut of overflow
  ManureDensities     <- calcOutput("CarbonManure", aggregate=FALSE)[,,"c"]
  Landuse             <- calcOutput("Landuse", aggregate=FALSE)
  ManureSOCbudget     <- dimSums(Landuse[,getYears(ManureDensities),"crop"]*ManureDensities, dim=1)

  print(mbind(dimSums(ManureMoinput, dim=3), dimSums(ManureSOCbudget, dim=3), dimSums(ManureSOCbudget, dim=3)/dimSums(ManureMoinput, dim=3)))

  ManureCorrected     <- ManureSOCbudget

  ManureExcreted      <- dimSums(calcOutput("Excretion", cellular = TRUE, attributes = "npkc", aggregate = FALSE)[,,"c"][,,c("confinement","stubble_grazing")], dim=c(1,3.3))

  print(mbind(dimSums(ManureExcreted, dim=3), dimSums(ManureMoinput, dim=3), dimSums(ManureCorrected, dim=3)))
  print(quantile(ManureApplication+ManureGrazing, probs=seq(0,1,0.02)))
  range(ManureApplication+ManureGrazing)
```

Same problem here. Manure share is going down.

```{r production}

#Production     <- collapseNames(calcOutput("Production", products="kcr", cellular=TRUE, attributes="dm", irrigation=FALSE, aggregate = FALSE))
Balance        <- calcOutput("FAOmassbalance", aggregate=FALSE)

CropProduction <- dimSums(Balance[,,findset("kcr")][,,"production"][,,"c"], dim=c(1,3))
FeedProduction <- dimSums(Balance[,,c("pasture",findset("kap"),findset("kres"),"scp"), invert=TRUE][,,"feed"][,,"c"], dim=c(1,3))
ResidueFeed    <- dimSums(Balance[,,findset("kres")][,,"feed"][,,"c"], dim=c(1,3))

print(mbind(CropProduction,FeedProduction,ResidueFeed))
```

Biomass excreted ist bigger than feed intake via plant harvest and residues taken for feed.

```{r soc}

  Litter             <- dimSums(calcOutput("CarbonLitter", aggregate=FALSE)[,,"c"]*Landuse[,,"natveg"], dim=1)

  # Calculate effective input
  param <- readSource("IPCCSoil", convert=FALSE)
  param.f1_metab2a          <- setYears(param[,,"f1"], NULL) 
  # stabilization efficiencies for metabolic decay products entering the active pool
  param.f3_struc2s          <- setYears(param[,,"f3"], NULL) 
  # stabilization efficiencies for structural decay products entering the slow pool
  param.f2_struc2a <- mbind(setNames(setYears(param[,,"f2_ft"], NULL), "crop"),
                            setNames(setYears(param[,,"f2_nt"], NULL), "natveg"))
  # stabilization efficiencies for structural decay products entering the active pool (tillage and no tillage specific)

  cell.input       <- mbind(add_dimension(dimSums(calcOutput("CarbonInput", landtype="crop", aggregate = FALSE),dim=3.2),
                                          dim=3.1, nm = "crop",   add="landuse"),
                            add_dimension(collapseNames(calcOutput("CarbonInput", landtype="natveg", aggregate = FALSE)),
                                          dim=3.1, nm = "natveg", add="landuse"))
  CarbonInput      <- dimSums(cell.input[,,"totalC"]*Landuse, dim=c(1))

  CarbonInputEff   <- collapseNames(dimSums((cell.input[,,"metabDOC"] * param.f1_metab2a +
                                             cell.input[,,"strucDOC"] * param.f2_struc2a +
                                             cell.input[,,"ligninC"]  * param.f3_struc2s) * Landuse, dim=c(1)))

  CarbonInputTot   <- collapseNames(dimSums(SOC[,,"steadystate"]*SOC[,,"decayrate"]*Landuse, dim=c(1,3.3)))

  # Load inputs, active steady state stock and decay rate

  SOC                <- calcOutput("SoilCarbon", output="full", aggregate=FALSE)
  Landuse            <- calcOutput("Landuse",       aggregate=FALSE)
  LanduseChange      <- calcOutput("LanduseChange", aggregate=FALSE)

  CarbonStocks       <- collapseNames(dimSums(SOC[,,c("steadystate","actualstate")]*Landuse, dim=1))
  CarbonTransfer     <- dimSums(toolConditionalReplace(SOC[,,"carbontransfer"], "<0", 0),dim=c(1,3.3))
  CarbonRelease      <- collapseNames(dimSums(SOC[,,"interstate"]*SOC[,,"decayrate"]*Landuse, dim=c(1, 3.3)))

print(mbind(CarbonTransfer, CarbonRelease))
```