### 2020-05-03 ResultNotebooks/scf_colormaps.Rmd

---
  title: "Stock Change Factor maps (coloured)"
output:
  html_document:
  df_print: paged
pdf_document: default
---

```{r load data}
library(moinput)

SoilCarbon <- readRDS("/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43_SOC_0.5.0/SoilCarbon.rds")
Landuse    <- readRDS("/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43_SOC_0.5.0/Landuse.rds")

setConfig(cachefolder  = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43/", forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(mappingfolder = "/home/kristine/mnt/rd3mod/inputdata/mappings/")

koeppen           <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )
getNames(koeppen) <- tolower(getNames(koeppen))
koepp2IPCC        <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
climateIPCC       <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
climateIPCC       <- climateIPCC[,getYears(SoilCarbon),]

```

```{r calculate SCF}

# one colour scheme
cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/
                  dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3)
cropland_share[Landuse[,,"crop"]==0] <- NA

# 4 colour scheme
test          <- cropland_share
test[Landuse[,,"crop"]==0] <- NA
test[test>2]  <- 1.9999
test[test==0] <- 0.0001
test          <- test*climateIPCC[,,1]+(test+2)*climateIPCC[,,2]+(test+4)*climateIPCC[,,3]+(test+6)*climateIPCC[,,4]

```

```{r plot density maps}

# No climate zone specification
library(RColorBrewer)
cols      <- colorRampPalette(brewer.pal(11, "RdYlGn"))(1000)

luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
png("CShare_1965.png",width = 1280, height = 720, res = 108)
luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
dev.off()

luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
png("CShare_2010.png",width = 1280, height = 720, res = 108)
luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
dev.off()

```

```{r plot density maps 4 colours}

# With climate zone specification
library(RColorBrewer)
pal_pink   <- c("#330022","#aa99aa","#ccbbcc","#FFBBEE")
pal_green  <- c("#003300","#99aa99","#bbccbb","#CCFFCC")
pal_blue   <- c("#000044","#9999aa","#bbbbcc","#CCFFFF")
pal_red    <- c("#FFEECC","#ccc2bb","#aaa299","#440000")

pal_pink   <- c("#110044","#774499","#aaaaaa","#FF0088","#FFAADD")
pal_green  <- c("#004400","#00FF00","#aaaaaa","#FFFF00","#FFFFCC")
pal_blue   <- c("#000044","#0000FF","#aaaaaa","#00FFFF","#CCFFFF")
pal_red    <- c("#FFDD99","#feb24c","#aaaaaa","#FF0000","#440000")

cols <- c(colorRampPalette(pal_pink)(500)[500:1],
          colorRampPalette(pal_red)(500),
          colorRampPalette(pal_blue)(500)[500:1],
          colorRampPalette(pal_green)(500)[500:1]
)

luplot::plotmap(test[,"y2010",], legend_range = c(0,8), legend_colours = cols, main="Stock Change Factor on Cropland 2010 (climate zones color coded)")
png("CShare_4ColorClimates.png", width = 1280, height = 720, res = 108)

luplot::plotmap(test[,"y2010",], legend_range = c(0,8), legend_colours = cols, main="Stock Change Factor on Cropland 2010 (climate zones color coded)")
dev.off()
```