### 2020-06-26 SOCbudgetPaper/appendix.Rmd
update method part + smaller stuff in other section and result figure

* table on method subsections to functions within R packages

* table on mapping LUH2FAO2MAG

* kcr2kres mapping

* carbon 2 dry matter @Ma2018 (??) -> Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation

* closs in AWMS - Table

* map on climate zone used for SCF