### 2020-08-14 SOCbudgetPaper/appendix.Rmd
update written part to first official version (woohoo)

# Appendix

## table on method subsections to functions within R packages {#append:subsection2mrfunctions}

## table on mapping LUH2FAO2MAG {#append:Tableluh2fao2mag}

## kcr2kres mapping {#append:Tablekcr2kres}

## carbon 2 dry matter {#append:Tablec2dm}
Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation

## closs in AWMS - Table {#append:TableclossAWMS}

## map on climate zone used for SCF {#append:climatemap}