### 2020-08-14 SOCbudgetPaper/conclusion.Rmd

\conclusions

<!-- You can modify the section name with  `\conclusions[modified heading if necessary]`. -->

Outlook including perspective on mitigation (and soc enhancement in the future).