### 2020-12-10 ResultNotebooks/results_SOCcompare.Rmd
mayor update on figures and flowfig

---
title: "SOC budget analysis in comparsion"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r load}
library(mrcommons)
library(mrsoil)
library(magpiesets)

rev       <- "rev0.11"
outputdir <- "Output/"

setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
#untar(paste0(getConfig("outputfolder"),"/",rev,"histManagement_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")

SoilCarbon <- read.magpie("Data/histManagement/SoilCarbon.rds")
Landuse    <- read.magpie("Data/histManagement/Landuse.rds")

```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

SOCStock     <- dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=3)
glo_SOCStock <- dimSums(SOCStock, dim=1) 
cell2iso     <- toolGetMapping("CountryToCellMapping.csv", type = "cell")
iso_SOCStock <- toolCountryFill(toolAggregate(SOCStock, rel=cell2iso, from="celliso", to="iso", dim=1), fill=0)
iso2reg      <- read.csv2("Data/histManagement/regionmappingH12.csv")
reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")
```

Load validation data:

```{r Validation Stocks}

library(mrvalidation)

  LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
  WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
  GSOC       <- calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate="REG+GLO", try=TRUE)
  SoilGrids  <- calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate="REG+GLO", try=TRUE)
  LPJmL4     <- calcOutput("ValidCarbonStocks", datasource = "LPJmL4Paper", aggregate="REG+GLO", try=TRUE)

```

```{r table glo soc stocks}

library(knitr)
library(stats)
library(luplot)
library(reshape2)

SOCStock <- mbind(glo_SOCStock, reg_SOCStock)
glo_SOC  <- mbind(setNames(SOCStock[,"y2010",],"SOC Budget"), 
                  setNames(LPJmL4[,"y2010",], "LPJmL4Paper"),
                  setYears(setNames(WISE[,1,],"WISE"), "y2010"),
                  setYears(setNames(SoilGrids[,1,],"SoilGrids"),"y2010"),
                  setYears(setNames(GSOC[,1,],"GSOC"),"y2010"))

#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))

comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(glo_SOC/1000)), 
                           Data1 ~ Region, value.var="Value", 
                            row.names=levels(as.data.frame(glo_SOC)$Data1)))

saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"gloSOC_comparison.rds"))

```

```{r kable2image}
library(kableExtra)

table <- readRDS("Output/gloSOC_comparison.rds")
colnames(table[[1]])[1] <- "Model/Database"
colnames(table[[1]])[5] <- "Global SOC in GtC"
save_kable(kable_styling(kable(table$comparison_table[c(1,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
         file="Output/Images/TableSOC_comparison.png")

save_kable(kable_styling(kable(table$comparison_table[setdiff(1:13,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
         file="Output/Images/TableRegSOC_comparison.png")

```

Plot regional stocks

```{r point plot reg stock, out.width=800, out.height=600}
library(ggplot2)

supp.labs <- getNames(glo_SOC["GLO",,invert=TRUE])
names(supp.labs) <- c("OJ", "VC")

leg_name <- "Data Source"

glob <- ggplot(data = as.data.frame(glo_SOC["GLO",,][,"y2010",]/1000), aes(Year,Value, group=Data1)) + 
  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
  scale_size_manual(values=c(4,2,2,2,2)) + 
  scale_shape_manual(values=c(19,6,2,5,0)) +
  labs(color  = leg_name, size = leg_name, shape = leg_name) +
  xlab("") + ylab("SOC (0-30cm) in GtC")

comp <- ggplot(data = as.data.frame(glo_SOC["GLO",,invert=TRUE][,"y2010",]/1000), aes(Year,Value, group=Data1)) + 
  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
  facet_wrap(~ Region, scales="free", drop=TRUE) + 
  scale_size_manual(values=c(4,2,2,2,2)) + 
  scale_shape_manual(values=c(19,6,2,5,0)) +
  labs(color  = leg_name, size = leg_name, shape = leg_name) +
  xlab("") + ylab("SOC (0-30cm) in GtC")

glob
comp

ggsave(paste0(outputdir,"/glo_comparisonfigure.png"), plot=glob, height=6, width=3, units="cm")
ggsave(paste0(outputdir,"/reg_comparisonfigure.png"), plot=comp)

```

```{r point plot glo stock, fig.width=3, fig.height=4}
library(ggplot2)

glob
ggsave(paste0(outputdir,"/glo_comparisonfigure.png"), plot=glob)
```
`