### 2020-12-10 ResultNotebooks/results_SOCtable.Rmd

---
title: "Table on stock change factors per clinate zone"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrsoil)
library(magpiesets)

rev       <- "rev0.11"
outputdir <- "Output/"

setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")

SoilCarbon <- read.magpie(paste0("Data/histManagement/SoilCarbon.rds"))[,1975:2010,]
Landuse    <- read.magpie(paste0("Data/histManagement/Landuse.rds"))[,1975:2010,]

```

```{r table scf climatezones}

library(knitr)
library(stats)
library(reshape2)

years             <- intersect(findset("past"), getYears(SoilCarbon))

### calculating climate mask
koeppen           <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )[,years,]
koepp2IPCC        <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
getNames(koeppen)         <- tolower(getNames(koeppen))
koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
climateIPCC       <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
rm(koeppen)

### calculating stock change factors
SOCDens           <- dimSums(SoilCarbon[,years,c("actualstate","naturalstate")], dim=3.3) 
SOCDens[,,"naturalstate.crop"] <- SOCDens[,,"actualstate.crop"]

CShare_ipcczone   <- dimSums(climateIPCC * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
                     dimSums(climateIPCC * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)

getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))

SCF_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,years,"naturalstate"])), 
                             Year ~ Data1, value.var="Value", row.names=years), row.names = "Year")
saveRDS(list(SCF_table = SCF_table), paste0(outputdir, "SFC_climatezone.rds"))

knitr::kable(SCF_table, "latex",digits=2, label="naturalstate_table")

```

```{r plot climte map}

magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
climate2plot <- dimSums(climateIPCC*new.magpie("GLO",NULL,getNames(climateIPCC),c(1:4)), dim=3)
climate2plot <- merge(as.data.frame(climate2plot), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")

library("ggplot2")
theme_set(theme_bw())
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library(RColorBrewer)
library(sp)
library(rworldmap)
library(oce) 
library(scales)

legend_size <- 10

map.world <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)  
coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])

cols          <- colorRampPalette(brewer.pal(4, "RdYlBu"))(4)

ClimateZones <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(climate2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = factor(Value))) +
    scale_fill_manual(name = expression("climate zones"), 
                      values = cols, na.value = "white",
                      breaks= c(1,2,3,4),
                      labels = gsub("_", " ", getNames(climateIPCC)),
                      guide = guide_legend(
                        title.theme=element_text(size=legend_size),
                        direction = "horizontal",
                        title.position = 'top',
                        # I shift the labels around, the should be placed 
                        # exactly at the right end of each legend key
                        title.hjust = 0.5,
                        label.hjust = 1,
                        nrow = 1,
                        byrow = F,
                        label.position = "bottom",
                        keyheight = 0.5,
                        keywidth  = 2
                      )) +
    theme(legend.position="bottom",
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          legend.text=element_text(size=legend_size),
          rect=element_blank(),
          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )
ClimateZones 
ggsave(paste0(outputdir,"/climatezones.png"), plot=ClimateZones , width = 10, height = 7, dpi = 300, units = "cm")
```

```{r table ipcc scf}

library(knitr)
library(stats)
library(luplot)
library(reshape2)

MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL,outer(c("IPCC2006","IPCC2019"),
                                                                          unique(koepp2IPCC$ipcc_reduced), paste, sep="."))

MShare_ipcc <- add_dimension(MShare_ipcc, dim=3.2, nm="medium.invariant", add="input.year")
LShare_ipcc <- add_dimension(LShare_ipcc, dim=3.2, nm="low.invariant",    add="input.year")
HShare_ipcc <- add_dimension(HShare_ipcc, dim=3.2, nm="high.invariant",   add="input.year")

MShare_ipcc[,,"IPCC2006.medium"][,,"tropical_moist"]  <- "0.48"
MShare_ipcc[,,"IPCC2006.medium"][,,"tropical_dry"]    <- "0.58--0.64" 
MShare_ipcc[,,"IPCC2006.medium"][,,"temperate_moist"] <- "0.69"
MShare_ipcc[,,"IPCC2006.medium"][,,"temperate_dry"]   <- "0.80"

MShare_ipcc[,,"IPCC2019.medium"][,,"tropical_moist"]  <- "0.83"
MShare_ipcc[,,"IPCC2019.medium"][,,"tropical_dry"]    <- "0.92" 
MShare_ipcc[,,"IPCC2019.medium"][,,"temperate_moist"] <- "0.69--0.70"
MShare_ipcc[,,"IPCC2019.medium"][,,"temperate_dry"]   <- "0.76--0.77"

#Low input regime factor
LShare_ipcc[,,"IPCC2006.low"][,,"tropical_moist"]  <- "0.44"
LShare_ipcc[,,"IPCC2006.low"][,,"tropical_dry"]    <- "0.55--0.61" 
LShare_ipcc[,,"IPCC2006.low"][,,"temperate_moist"] <- "0.66"
LShare_ipcc[,,"IPCC2006.low"][,,"temperate_dry"]   <- "0.74"

LShare_ipcc[,,"IPCC2019.low"][,,"tropical_moist"]  <- "0.76"
LShare_ipcc[,,"IPCC2019.low"][,,"tropical_dry"]    <- "0.87" 
LShare_ipcc[,,"IPCC2019.low"][,,"temperate_moist"] <- "0.66--0.67"
LShare_ipcc[,,"IPCC2019.low"][,,"temperate_dry"]   <- "0.70--0.71"

#High input regime factor
HShare_ipcc[,,"IPCC2006.high"][,,"tropical_moist"]  <- "0.53"
HShare_ipcc[,,"IPCC2006.high"][,,"tropical_dry"]    <- "0.60--0.67" 
HShare_ipcc[,,"IPCC2006.high"][,,"temperate_moist"] <- "0.77"
HShare_ipcc[,,"IPCC2006.high"][,,"temperate_dry"]   <- "0.83"

HShare_ipcc[,,"IPCC2019.high"][,,"tropical_moist"]  <- "0.92"
HShare_ipcc[,,"IPCC2019.high"][,,"tropical_dry"]    <- "0.96" 
HShare_ipcc[,,"IPCC2019.high"][,,"temperate_moist"] <- "0.77--0.78"
HShare_ipcc[,,"IPCC2019.high"][,,"temperate_dry"]   <- "0.79--0.80"

CShare_ipcc <- mbind(LShare_ipcc,MShare_ipcc,HShare_ipcc)
getNames(CShare_ipcc) <- gsub("_", " ", getNames(CShare_ipcc))

comparison <- mbind(CShare_ipcc, 
                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1990","naturalstate"]),NULL),
                                  nm="This Study.hist.1990",add="model.input.year"),2),
                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),NULL),
                                  nm="This Study.hist.2010",add="model.input.year"),2))

#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))

comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(comparison)), 
                           Data1 + Data2 + Data3 ~ Data4, value.var="Value", 
                            row.names=levels(as.data.frame(comparison)$Data1)))

colnames(comparison_table) <- gsub("\\.", " ", colnames(comparison_table))
colnames(comparison_table)[1:3] <- c("Source", "Input", "Year")

saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"SFC_comparison.rds"))

```

```{r kable2image}
library(kableExtra)

getwd()
table <- readRDS(paste0(outputdir,"/SFC_comparison.rds"))

save  <- kbl(table$comparison_table, latex = T, digit=2)%>%
            kable_styling(latex_options = c("scale_down","striped"), full_width = F)%>%
              row_spec(7:8, bold=TRUE)

save_kable(save, file="Output/Images/TableSCF_comparison.png")
```