### 2020-12-14 ResultNotebooks/results_SOCcompare.Rmd

---
title: "SOC budget analysis in comparsion"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r load}
library(magclass)
library(madrat)

outputdir <- "Output/"

SoilCarbon <- read.magpie("Data/histManagement/SoilCarbon.rds")
Landuse    <- read.magpie("Data/histManagement/Landuse.rds")

```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

SoilCarbon[,,c("naturalstate.crop")] <- SoilCarbon[,,c("naturalstate.natveg")]
SOCStock     <- dimSums(dimSums(SoilCarbon[,,c("naturalstate","actualstate")], dim=3.3)* Landuse, dim=3.2)
cell2reg     <- read.csv("Data/histManagement/ClimateZoneMapping.csv")
reg_SOCStock <- toolAggregate(SOCStock, rel=cell2reg, from="cell", to="IPCC")
```

```{r table glo soc stocks}

  WISE       <- readRDS("Data/histManagement/WISE.rds")
  GSOC       <- readRDS("Data/histManagement/GSOC.rds")
  SoilGrids  <- readRDS("Data/histManagement/SoilGrids.rds")
  LPJmL4     <- readRDS("Data/histManagement/LPJmL4.rds")

library(knitr)
library(stats)
library(reshape2)

SOCStock <- reg_SOCStock
reg_SOC  <- mbind(setNames(SOCStock[,"y2010","actualstate"],"This Study"),
                  setNames(SOCStock[,"y2010","naturalstate"],"This Study (PNV)"),
                  setNames(LPJmL4[,"y2010",], "LPJmL4Paper"),
                  setYears(setNames(WISE[,1,],"WISE"), "y2010"),
                  setYears(setNames(SoilGrids[,1,],"SoilGrids"),"y2010"),
                  setYears(setNames(GSOC[,1,],"GSOC"),"y2010"))
glo_SOC <- mbind(setCells(dimSums(reg_SOC, dim=1),"GLO"), reg_SOC)

#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))

comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(glo_SOC/1000)), 
                           Data1 ~ Region, value.var="Value", 
                            row.names=levels(as.data.frame(glo_SOC)$Data1)))

comparison_table$Year <- "2010"

saveRDS(object = comparison_table, paste0(outputdir,"gloSOC_comparison.rds"))

```

```{r point plot reg stock, out.width=800, out.height=600}
library(ggplot2)
library(knitr)
library(stats)
library(reshape2)

table    <- readRDS("Output/gloSOC_comparison.rds")
leg_name <- ""

names(table) <- c("Data1","Boreal Dry", "Boreal Moist", "C-Temp Dry", "C-Temp Moist","GLO","Polar Dry", "Polar Moist",
                  "Tropical Dry", "Tropical Moist", "Tropical Montane", "Tropical Wet", "W-Temp Dry", "W-Temp Moist", "Year")
table        <- table[,c(1,15,6,7,8,2,3,4,5,13,14,9,10,11,12)]

cols   <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")[c(7,3,1,4,6,8)]

glob <- ggplot(data = melt(subset(table, select =c(Data1, GLO, Year))), aes(y=Year,x=value, group=Data1)) + 
  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
  scale_size_manual(values=c(4,2,2,2,2,2)) + 
  scale_shape_manual(values=c(19,16,6,2,5,0)) +
  scale_color_manual(values = cols) +
  labs(color  = leg_name, size = leg_name, shape = leg_name) +
  ylab("") + xlab("SOC (0-30cm) in GtC") +
  theme(axis.ticks.y =element_blank(),
        axis.text.y=element_blank(), 
        legend.position="bottom") + 
  guides(fill=guide_legend(nrow=2,byrow=TRUE))
        #legend.position="right") + guides(fill=guide_legend(ncol=1,byrow=TRUE))

comp <- ggplot(data = melt(subset(table, select=-GLO)), aes(y=Year,x=value, group=Data1)) + 
  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
  facet_wrap(~ variable, scales="free", drop=TRUE, labeller = label_wrap_gen(width=18)) + 
  scale_size_manual(values=c(4,2,2,2,2,2)) + 
  scale_shape_manual(values=c(19,16,6,2,5,0)) +
  scale_color_manual(values = cols) +
  labs(color  = leg_name, size = leg_name, shape = leg_name) +
  ylab("") + xlab("SOC (0-30cm) in GtC") +
  theme(axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        legend.position = "bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))

glob
comp

ggsave(paste0(outputdir,"/Images/glo_comparisonfigure.png"), plot=glob, height=4, width=10, units="cm", dpi = 300)
ggsave(paste0(outputdir,"/Images/reg_comparisonfigure.png"), plot=comp, height=9, width=15, units="cm", dpi = 300)

```

```{r point plot glo stock, fig.width=3, fig.height=4}
library(ggplot2)

glob
ggsave(paste0(outputdir,"/glo_comparisonfigure.png"), plot=glob)
```
`