### 2020-12-14 ResultNotebooks/results_SOCflows.Rmd
Flow figure update

---
  title: "First SOC budget flow analysis"
output:
  html_document:
  df_print: paged
pdf_document: default
---

```{r load}
library(magclass)
library(madrat)
library(magpiesets)

outputdir <- "Output/"
revfolder <- "Data/histManagement/"

```

```{r load data}

year  <- "y2010"
years <- c(paste0("y",as.numeric(substring("y2010",2))-1), year)

CarbonInput      <- readRDS(paste0(revfolder,"CarbonInput.rds"))[,year,]
CarbonLitter     <- readRDS(paste0(revfolder,"CarbonLitter.rds"))[,year,]
CarbonManure     <- readRDS(paste0(revfolder,"CarbonManure.rds"))[,year,]          
CarbonResidues   <- readRDS(paste0(revfolder,"CarbonResidues.rds"))[,year,]
CarbonHarvest    <- readRDS(paste0(revfolder,"Crop_Harvest.rds"))[,year,]
CarbonVegetation <- readRDS(paste0(revfolder,"CarbonVegetation.rds"))[,years,]
SoilCarbonOut    <- readRDS(paste0(revfolder,"SoilCarbon.rds"))[,years,]
CarbonInflow     <- readRDS(paste0(revfolder,"CarbonInflow.rds"))[,years,]
CarbonDecay      <- readRDS(paste0(revfolder,"Decay.rds"))[,year,]

Landuse          <- readRDS(paste0(revfolder,"Landuse.rds"))[,years,]
LanduseChange    <- readRDS(paste0(revfolder,"LanduseChange.rds"))[,years,]

FAOmassbalance   <- readRDS(paste0(revfolder,"FAOmassbalance_ISO.rds"))[,year,"c"]
ManureExcreted   <- readRDS(paste0(revfolder,"Manure_excreted.rds"))[,year,"c"]
ManureRecycled   <- readRDS(paste0(revfolder,"Manure_recycled.rds"))[,year,"c"]
ResidueBiomass   <- readRDS(paste0(revfolder,"Residue_Biomass.rds"))[,year,"c"]
ResiduesAgDemand <- readRDS(paste0(revfolder,"ResiduesAg_FieldBalance.rds"))[,year,"c"]

grepSpaceTimeC    <- function(x){return(collapseNames(dimSums(x[,year,"c"],dim=1)))}
grepSpaceTime     <- function(x){return(collapseNames(dimSums(x[,year,],dim=1)))}

TotalSOCBudget    <- NULL

#Agriculture
kcr        <- findset("kcr")
kli        <- findset("kli")
kres       <- findset("kres")
kExResPast <- setdiff(findset("kall"),c(kres,kli,"pasture","wood","woodfuel","scp","fish"))
kPrim      <- setdiff(findset("kall"),c(kli,"scp","fish"))

CropHANPP         <- dimSums(grepSpaceTimeC(FAOmassbalance[,,kcr][,,c("food","waste","other_util","bioenergy","feed",
                                                                     "milling","refining","extracting","fermentation","distilling")]), dim=3)

CropFEED          <- grepSpaceTimeC(dimSums(FAOmassbalance[,,kExResPast][,,"feed"], dim=3.1))
CropPRIM          <- CropHANPP-CropFEED

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(CropFEED, "crop_feedintake"),
                           setNames(CropPRIM, "crop_humandemand"))

ResFEED           <- grepSpaceTimeC(dimSums(FAOmassbalance[,,kres][,,"feed"], dim=3.1))
ResPRIM           <- dimSums(grepSpaceTimeC(FAOmassbalance[,,kres][,,c("other_util","bioenergy")]), dim=3)

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(ResFEED, "ag_feedintake"),
                           setNames(ResPRIM, "ag_humandemand"))

FEED              <- grepSpaceTimeC(dimSums(FAOmassbalance[,,kPrim][,,"feed"], dim=3.1))
LivstHANPP        <- dimSums(grepSpaceTimeC(FAOmassbalance[,,kli][,,c("food","waste","other_util")]), dim=3)
LivstCropHANPP    <- (CropFEED+ResFEED)/FEED * LivstHANPP

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(LivstCropHANPP, "livst_humandemand"))

ResidueBiomassTot <- grepSpaceTimeC(dimSums(ResidueBiomass, dim=3.2))
ResidueRecycledTot<- grepSpaceTimeC(dimSums(ResiduesAgDemand[,,"recycle"], dim=3.2))
ResidueLossTot    <- ResidueBiomassTot[,,"ag"] - ResidueRecycledTot - ResFEED - ResPRIM

ResidueInput      <- grepSpaceTimeC(Landuse[,2,"crop"]* CarbonResidues)

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(ResidueBiomassTot[,,c("ag","bg")], c("ag_prod","bg_prod")),
                           #setNames(ResidueRecycledTot, "ag_recycle"),
                           #setNames(ResidueBiomassTot[,,"bg"], "bg_recycle"),
                           setNames(ResidueInput[,,c("ag","bg")], c("ag_recycle","bg_recycle")),
                           setNames(ResidueLossTot, "ag_loss"))

ManureExcretedTot <- grepSpaceTimeC(dimSums(ManureExcreted[,,c("stubble_grazing","confinement")], dim=c(3.1,3.2)))
#ManureRecycledTot <- grepSpaceTimeC(dimSums(ManureRecycled, dim=3.1))
ManureInput       <- dimSums(grepSpaceTimeC(Landuse[,2,"crop"]* CarbonManure), dim=3.1)
ManureLossTot     <- ManureExcretedTot - ManureInput

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(ManureExcretedTot, "manure_prod"),
                           setNames(ManureInput, "manure_recycle"),
                           setNames(ManureLossTot    , "manure_loss"))

LitterRecycleTot  <- grepSpaceTimeC(CarbonLitter)

CropBiomassTot    <- dimSums(ResidueBiomassTot, dim=3) + dimSums(grepSpaceTimeC(FAOmassbalance[,,kcr][,,"production"]),dim=3)

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(CropBiomassTot, "crop_prod"))

LivstCropLoss     <- CropFEED - ManureExcretedTot - LivstCropHANPP 
CropPRIMLoss      <- CropPRIM + ResPRIM + LivstCropHANPP - dimSums(grepSpaceTimeC(FAOmassbalance[,,c(kcr,kli,kres)][,,"other_util"]), dim=3) # only material use is not re-emitted

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(LivstCropLoss, "livst_loss"),
                           setNames(CropPRIMLoss,  "humandemand_loss"))

### Carbon To
CarbonTo            <- dimSums(grepSpaceTime(toolConditionalReplace(SoilCarbonOut[,,"carbontransfer"], "<0", 0)), dim=c(3.2))
CarbonToEff         <- dimSums(grepSpaceTime(SoilCarbonOut[,,"carbontransfer"]), dim=c(3.2))

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(CarbonTo[,,"natveg"], "to_natveg"),
                           setNames(CarbonTo[,,"crop"],   "from_natveg"))

SOCsub            <- grepSpaceTime(SoilCarbonOut[,,"actualstate"]*Landuse)
SOCtot            <- dimSums(SOCsub, dim=3.2)

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(SOCsub, gsub("\\.","_", getNames(SOCsub))),
                           setNames(SOCtot, paste0("soc_", getNames(SOCtot))))

CarbonIn          <- grepSpaceTime(CarbonInflow[,,"crop"]*Landuse[,,"crop"])
CarbonIn          <- mbind(CarbonIn[,,"ag"],CarbonIn[,,"bg"],
                           add_dimension(dimSums(
                             CarbonIn[,,c("livst_chick","livst_egg","livst_milk","livst_pig","livst_rum")], dim=3.1),
                              add="subpool", nm="manure"))

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(CarbonIn, gsub("\\.","_", getNames(CarbonIn))))
CarbonInTot       <- dimSums(CarbonIn, dim=3.2)

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(ManureInput - CarbonInTot[,,"manure"], "manure_decay"),
                           setNames(ResidueInput[,,"ag"] - CarbonInTot[,,"ag"], "ag_decay"),
                           setNames(ResidueInput[,,"bg"] - CarbonInTot[,,"bg"], "bg_decay"))

SOCIn             <- dimSums(CarbonInTot, dim=3)
SOCtransfer       <- CarbonToEff[,,"crop"] 
SOCPrev           <- dimSums(grepSpaceTime(setYears(
                       SoilCarbonOut[,1,"actualstate"][,,"crop"]*Landuse[,1,"crop"], year)), dim=3)
SOCLoss           <- SOCPrev - SOCtot[,,"crop"] + SOCtransfer + SOCIn

TotalSOCBudget    <- mbind(TotalSOCBudget,
                           setNames(SOCLoss, "soc_decay"))

write.magpie(TotalSOCBudget, paste0(outputdir,"TotalSOCBudget.cs3"))

```

```{r}

DecayCut          <- Decay
DecayCut[Decay>1] <- 1

### Carbon In
f1        <- param[,,"f1"]
f2_crop   <- param[,,"f2_ft"]
f2_natveg <- param[,,"f2_nt"]
f3        <- param[,,"f3"]

# f2 has to be updated with new test runs

MineralizeFracInput <- setCells(mbind(add_dimension(collapseNames(CarbonInput[,,"totalC",invert=TRUE][1,,"res"]), dim=3.1, add = "lu", nm = "crop"),
                                      add_dimension(collapseNames(CarbonInput[,,"totalC",invert=TRUE][1,,"res"]), dim=3.1, add = "lu", nm = "natveg")), "GLO")
MineralizeFracInput[,,"metabDOC"]        <- f1
MineralizeFracInput[,,"crop.strucDOC"]   <- f2_crop
MineralizeFracInput[,,"natveg.strucDOC"] <- f2_natveg
MineralizeFracInput[,,"ligninC"]         <- f3

input2lu <- as.data.frame(list(c("res", findset("kli"), "litfall"),
                            c(rep("crop",6), "natveg")), col.names=c("input","lu"))

CarbonInEff       <- collapseNames(dimSums(toolAggregate(CarbonInput[,"y2010",][,,"totalC",invert=TRUE], rel=input2lu, dim=3.1) *
                                            MineralizeFracInput[,"y2010",] * Landuse[,"y2010",], dim=c(1,3.2)))

CarbonInSource   <- collapseNames(dimSums(CarbonInput[,"y2010",][,,"totalC",invert=TRUE]  * 
                                            toolAggregate(MineralizeFracInput[,"y2010",] * Landuse[,"y2010",], rel=input2lu, from="lu", to="input", dim=3.1), dim=c(1,3.2)))

CarbonInRaw       <- collapseNames(dimSums(SoilCarbonOut[,"y2010","steadystate"]*DecayCut[,"y2010",]*Landuse[,"y2010",], dim=c(1,3.3)))

### Carbon Out

CarbonOutRaw      <- collapseNames(dimSums(SoilCarbonOut[,"y2010","interstate"]*DecayCut[,"y2010",]*Landuse[,"y2010",], dim=c(1,3.3)))
CarbonOutEff      <- CarbonOutRaw + CarbonInEff - CarbonInRaw

### SOC State

SOCstate <- dimSums(SoilCarbonOut[,c("y2009","y2010"),c("actualstate","interstate")]*Landuse[,c("y2009","y2010"),], dim=c(1,3.3))

```