### 2021-07-24 ResultNotebooks/create_LUtransition.Rmd
update results scripts

---
  title: "LU transitions overview figs"
output:
  md_document:
  variant: markdown_github
---

```{r arrange plot, out.height=1000 , out.width=2000}
library(cowplot)
library(ggplot2)

start     <- "../FlowFigure/LUC_start.pdf"
trans     <- "../FlowFigure/LUC_transition.pdf"
hetero    <- "../FlowFigure/LUC_hetero.pdf"
homo      <- "../FlowFigure/LUC_homo.pdf"
text      <- "../FlowFigure/LUC_text.pdf"
outputdir <- "Output/Images"

p1 <- ggdraw() + draw_image(start)
p2 <- ggdraw() + draw_image(trans)
p3 <- ggdraw() + draw_image(hetero)
p4 <- ggdraw() + draw_image(homo)
pt <- ggdraw() + draw_image(text)

grid_1 <- plot_grid(p1, p2, p3, ncol=3)
grid_2 <- plot_grid (pt, p4, ncol=2, rel_widths = c(2, 1))
grid   <- plot_grid(grid_1, grid_2, nrow=2)
grid
ggsave(paste0(outputdir,"/LUtransition.png"), plot=grid, width = 20, height = 10, dpi = 400, units = "cm", bg="white")

```