### 2021-07-24 ResultNotebooks/results_NC.Rmd

---
  title: "Global N release binding"
output:
  md_document:
  variant: markdown_github
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrsoil)
library(magpiesets)

rev       <- "rev19"
outputdir <- "Output/"

#filesNeeded  <- c("SoilCarbon.rds","Landuse.rds","LanduseChange.rds")
#untar(paste0(getConfig("outputfolder"),"/",rev, "histManagement_ResAllom_2021_01_12_h12_carbonbudget.tgz"), files=filesNeeded, exdir="Data/histManagement_ResAllom", extras="--keep-newer-files ")

# SoilCarbon    <- read.magpie(paste0("Data/histManagement_ResAllom/SoilCarbon.rds"))
# Landuse       <- read.magpie(paste0("Data/histManagement_ResAllom/Landuse.rds"))
# LanduseChange <- read.magpie(paste0("Data/histManagement_ResAllom/LanduseChange.rds"))

SoilCarbon    <- read.magpie(paste0("Data/rev19histManagement/SoilCarbon.rds"))
Landuse       <- read.magpie(paste0("Data/rev19histManagement/Landuse.rds"))
LanduseChange <- read.magpie(paste0("Data/rev19histManagement/LanduseChange.rds"))

```

Calc global numbers

```{r Carbon Emissions (glo)}

SoilCarbon        <- dimSums(SoilCarbon, dim="subpool")

# Carbon lost at Cropland staying Cropland

years     <- getYears(SoilCarbon, as.integer = TRUE)
#years     <- years[-1]

StayingCropland <- (SoilCarbon[,tail(years,-1),"actualstate.crop"]-setYears(SoilCarbon[,head(years,-1),"actualstate.crop"], head(years,-1)+1))*(Landuse[,tail(years,-1),"crop"]-LanduseChange[,tail(years,-1),"expansion.crop"])

# Carbon lost at newly comverted Cropland

NewlyConverted <- (SoilCarbon[,tail(years,-1),"actualstate.crop"]-setYears(SoilCarbon[,head(years,-1),"actualstate.natveg"], head(years,-1)+1))*(LanduseChange[,tail(years,-1),"expansion.crop"])

StayingCropland_gain <- StayingCropland_loss <- StayingCropland
StayingCropland_gain[StayingCropland<0] <- 0
StayingCropland_loss[StayingCropland>0] <- 0

NewlyConverted_gain <- NewlyConverted_loss <- NewlyConverted
NewlyConverted_gain[NewlyConverted<0] <- 0
NewlyConverted_loss[NewlyConverted>0] <- 0

NetChange <- NetChange_gain <- NetChange_loss <- NewlyConverted + StayingCropland
NetChange_gain[NetChange<0] <- 0
NetChange_loss[NetChange>0] <- 0

diffs <- mbind(setNames(NetChange,      "tot.net"),
               setNames(NetChange_gain, "tot.gain"),
               setNames(NetChange_loss, "tot.loss"),
               setNames(StayingCropland,      "stay.net"),
               setNames(StayingCropland_gain, "stay.gain"),
               setNames(StayingCropland_loss, "stay.loss"),
               setNames(NewlyConverted,       "conv.net"),
               setNames(NewlyConverted_gain,  "conv.gain"),
               setNames(NewlyConverted_loss,  "conv.loss")
               )

iso2cell <- toolGetMapping("CountryToCellMapping.csv", type="cell")
diffsIso <- toolAggregate(diffs, rel=iso2cell, from="celliso", to="iso", dim=1)

NChange <- dimSums(diffs, dim=1)/15
NChangeIso <- diffsIso/15

print(NChange)
print(NChangeIso[,"y2010",])
write.magpie(NChangeIso,"Output/NChange.csv")
write.magpie(NChange,"Output/NChangeGlo.csv")

```

```{r som}

SOM  <- read.magpie("Data/additional/SOM.nc")
SOM2 <- read.magpie("Data/additional/SOM2.nc")

years    <- getYears(SOM, as.integer = TRUE)
getCells(SOM) <- getCells(SOM2) <- getCells(Landuse)
getNames(SOM, dim=1) <- getNames(SOM2, dim=1) <- c("crop", "natveg")

SOC  <- mbind(add_dimension(collapseNames(toolConditionalReplace(SOM[,, "soilc"]/Landuse,"is.na()", 0)), dim=3.1, add="model", nm="som"), 
              add_dimension(collapseNames(toolConditionalReplace(SOM2[,,"soilc"]/Landuse,"is.na()", 0)), dim=3.1, add="model", nm="som2"))

StayingCroplandSOM <- collapseNames((SOC[,tail(years,-1),"crop"]-setYears(SOC[,head(years,-1),"crop"],head(years,-1)+1))*
  (Landuse[,tail(years,-1),"crop"]-LanduseChange[,tail(years,-1),"expansion.crop"]))

# Carbon lost at newly converted Cropland

NewlyConvertedSOM <- collapseNames((SOC[,tail(years,-1),"crop"]-setYears(SOC[,head(years,-1),"natveg"],head(years,-1)+1))*
  (LanduseChange[,tail(years,-1),"expansion.crop"]))

StayingCroplandSOM_gain <- StayingCroplandSOM_loss <- StayingCroplandSOM
StayingCroplandSOM_gain[StayingCroplandSOM<0] <- 0
StayingCroplandSOM_loss[StayingCroplandSOM>0] <- 0

NewlyConvertedSOM_gain <- NewlyConvertedSOM_loss <- NewlyConvertedSOM
NewlyConvertedSOM_gain[NewlyConvertedSOM<0] <- 0
NewlyConvertedSOM_loss[NewlyConvertedSOM>0] <- 0

NetChangeSOM <- NetChangeSOM_gain <- NetChangeSOM_loss <- NewlyConvertedSOM + StayingCroplandSOM
NetChangeSOM_gain[NetChangeSOM<0] <- 0
NetChangeSOM_loss[NetChangeSOM>0] <- 0

diffsom <- mbind(add_dimension(NetChangeSOM, dim=3.2, add="land.mode", nm="tot.net"),
                 add_dimension(NetChangeSOM_gain, dim=3.2, add="land.mode", nm="tot.gain"),
                 add_dimension(NetChangeSOM_loss, dim=3.2, add="land.mode", nm="tot.loss"),
                 add_dimension(StayingCroplandSOM,      dim=3.2, add="land.mode", nm="stay.net"),
                 add_dimension(StayingCroplandSOM_gain, dim=3.2, add="land.mode", nm="stay.gain"),
                 add_dimension(StayingCroplandSOM_loss, dim=3.2, add="land.mode", nm="stay.loss"),
                 add_dimension(NewlyConvertedSOM,       dim=3.2, add="land.mode", nm="conv.net"),
                 add_dimension(NewlyConvertedSOM_gain,  dim=3.2, add="land.mode", nm="conv.gain"),
                 add_dimension(NewlyConvertedSOM_loss,  dim=3.2, add="land.mode", nm="conv.loss")
)

iso2cell <- toolGetMapping("CountryToCellMapping.csv", type="cell")
diffsomIso <- toolAggregate(diffsom, rel=iso2cell, from="celliso", to="iso", dim=1)

NChangeSOM <- dimSums(diffsom, dim=1)/15
NChangeSOMIso <- diffsomIso/15

print(NChangeSOM)
print(NChangeSOMIso[,"y2010",])

NChangeALL    <- mbind(NChangeSOM, add_dimension(NChange, dim=3.1, add="model", nm="mrsoil"))
NChangeIsoALL <- mbind(NChangeSOMIso, add_dimension(NChangeIso, dim=3.1, add="model", nm="mrsoil"))

write.magpie(NChangeSOMIso,"Output/NChangeALL.csv")
write.magpie(NChangeSOM,"Output/NChangeALLGlo.csv")

SOMloss <- dimSums(-SOM[,,"delta_soilc"][,,"crop"]/15, dim=1)

```