### 2021-07-24 ResultNotebooks/results_SOCscenario.Rmd

---
title: "Global SOC scenarios"
output:
  md_document:
    variant: markdown_github
  html_document:  
    pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrsoil)
library(magpiesets)

rev         <- "rev26"
scenarios   <- c("rev26histManagement", "rev26constManagement", "rev26constManagement2", "rev26constManagement_constClimate", "rev26constManagement2_constClimate", "rev26constManure", "rev26constResidues", "rev26constResrate", "rev26constTillage", "rev26constYield", "rev26Initial-1901_constManagement", "rev26Initial-1901_constManagement2", "rev26Initial-1901", "rev26Initial-0_constManagement", "rev26Initial-0_constManagement2", "rev26Initial-0")

# filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
# setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
# 
# untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
#       exdir=paste0("Data/",rev,"histManagement"), extras="--keep-newer-files ")

# for(item in scenarios){
#    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
#    files=filesNeeded, exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
# }

outputdir <- "Output/"

```

```{r Load data and calc globals}

glo_SOCStock <- NULL

for(scen in scenarios){
  SoilCarbon <- read.magpie(paste0("Data/", scen, "/SoilCarbon.rds"))
  Landuse    <- read.magpie(paste0("Data/", scen, "/Landuse.rds"))
  SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
  SOCStock        <- toolConditionalReplace(dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2), "is.nan()",0)
  glo_SOCStock    <- mbind(glo_SOCStock, add_dimension(dimSums(SOCStock, dim=1), 
                                                       dim=3.1, nm=scen, add="scenario")) 
}

glo_SOCDebt  <- collapseNames(glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
glo_SOCShare <- collapseNames(toolConditionalReplace(glo_SOCStock[,,"actualstate"]/glo_SOCStock[,,"naturalstate"], "is.na()",0) )

saveRDS(mbind(add_dimension(collapseNames(glo_SOCStock[,,"actualstate"]), dim=3.1, add="type", nm="stock"),
              add_dimension(              glo_SOCDebt,                    dim=3.1, add="type", nm="debt"),
              add_dimension(              glo_SOCShare,                   dim=3.1, add="type", nm="share")), 
        paste0(outputdir, "glo_SOCscenarions.rds"))
```

```{r Carbon inputs (glo)}

CarbonInput  <- read.magpie(paste0("Data/rev26histManagement/CarbonInput.rds"))
Landuse      <- read.magpie(paste0("Data/rev26histManagement/Landuse.rds"))
glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))/collapseNames(dimSums(Landuse, dim=c(1)))

ci2plot <- c('above-ground residues'="ag.crop",          
             'below-ground residues'="bg.crop",          
             'manure from monogastrics'="livst_chick.crop",
             'manure from monogastrics'="livst_egg.crop",
             'manure from monogastrics'="livst_pig.crop",
             'manure from ruminants'="livst_milk.crop",  
             'manure from ruminants'="livst_rum.crop",   
             'natural litterfall wood fraction'="wood.natveg",
             'natural litterfall leaf fraction'="leaf.natveg")

glo_SOCInput <- toolAggregate(glo_SOCInput, rel=data.frame(from=ci2plot,to=names(ci2plot)), from="from", to="to", dim=3)

```

```{r Plot Management Scenarios}

glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])

manage_scen <-  c(histManagement   = "rev26histManagement",
                  constTillage     = "rev26constTillage", 
                  constManure      = "rev26constManure",
                  constResidues    = "rev26constResidues", 
                  constManagement  = "rev26constManagement")#,
                  #constManagementNoCC = "rev26constManagement_constClimate")

library(ggplot2)
library(scales)
library(gridExtra)
library(cowplot)

theme_set(theme_bw())
legend_size <- 10
scen_cols   <- c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000")[5:1]
input_cols  <- scen_cols[1:4]

limits <- c(-40,-32)
breaks <- c(-32, -34,- 36, -38, -40)

hSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),manage_scen]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
  labs( x = "Year", 
        y = expression(paste('Global \u0394SOC in GtC')),
        color="") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous(limits=limits, breaks=breaks) +
  scale_color_manual(values=scen_cols, labels = names(manage_scen)) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
        legend.position="bottom", legend.direction="vertical",
        legend.margin=margin(-10,-10,0,0),
        legend.box.margin=margin(-10,-10,0,0))

hSOCinput <- ggplot(data= as.data.frame(glo_SOCInput[,seq(1975,2010,1),c(3:4,2,1)]),  aes(x = Year, y = Value, group=Data1, fill=Data1)) + geom_area() +
  labs( x = "Year", 
        y = "SOC input in tC per ha\n (global average)",
        fill="") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous() +
  scale_fill_manual(values=input_cols) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
        legend.position="bottom", 
        legend.direction="vertical",
        legend.margin=margin(-10,-10,0,0),
        legend.box.margin=margin(-10,-10,0,0))

vSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1974,2010,1),manage_scen]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
  labs( x = "Year", 
        y = expression(paste('Global \u0394SOC in GtC')),
        color="") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous(limits = limits,
                     breaks = breaks) +
  scale_color_manual(values=scen_cols, labels = names(manage_scen)) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
        legend.position="right", legend.direction="vertical")

vSOCinput <- ggplot(data= as.data.frame(glo_SOCInput[,seq(1975,2010,1),c(3:4,2,1)]),  aes(x = Year, y = Value, group=Data1, fill=Data1)) + geom_area() +
  labs( x = "Year", 
        y = "SOC input in tC per ha\n (global average)",
        fill="") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous() +
  scale_fill_manual(values=input_cols) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size), legend.direction="vertical")

hSOCemis
hSOCinput
vSOCemis
vSOCinput

hgrid <- plot_grid(hSOCemis, hSOCinput, ncol=2, labels=c("(a)","(b)"),
                  label_fontface = "plain", label_size = 10, align="hv")

vgrid <- plot_grid(plot_grid(vSOCemis  + theme(legend.position = "none") , 
                             vSOCinput + theme(legend.position = "none"), ncol=1, labels=c("(a)","(b)"),
                             label_fontface = "plain", label_size = 10, align="v", axis="tblr"),
                   plot_grid(get_legend(vSOCemis), get_legend(vSOCinput), ncol=1,
                             label_fontface = "plain", label_size = 10, align="v"),
                   align="h")

hgrid
vgrid

ggsave(paste0(outputdir,"Images/scenario_horiz.png"), plot=hgrid, width = 15, height = 10, dpi = 300, units = "cm")
ggsave(paste0(outputdir,"Images/scenario_verti.png"), plot=vgrid, width = 12, height = 15, dpi = 300, units = "cm")

```

```{r Plot Initialization Scenarios}

glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])

init_scen   <-  c(`histManagement.Initial-natveg`          = "rev26Initial-1901", 
                  `constManagement.Initial-natveg`         = "rev26Initial-1901_constManagement",
                  `histManagement.Initial-eq`              = "rev26Initial-0", 
                  `constManagement.Initial-eq`             = "rev26Initial-0_constManagement",
                  `histManagement.Initial-spinup1530`      = "rev26histManagement", 
                  `constManagement.Initial-spinup1530`     = "rev26constManagement")

glo_SOCDebt <- glo_SOCDebt[,,init_scen]
getNames(glo_SOCDebt) <- names(init_scen)

library(ggplot2)
library(scales)
library(gridExtra)
library(cowplot)

theme_set(theme_bw())
legend_size <- 10
scen_cols   <- c("#648FFF","#785EF0","#DC267F")
input_cols  <- scen_cols[1:4]

limits <- c(-55,-25)
breaks <- c( -30,-40,-50 )

initSOCemis <- ggplot() + geom_path(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),]), 
                                    aes(x = Year, y = Value, group=interaction(Data1,Data2), linetype=Data1, color=Data2)) +
  labs( x = "Year", 
        y = expression(paste('Global \u0394SOC in GtC')),
        color    ="Initialization",
        linetype ="Management") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous(limits=limits, breaks=breaks) +
  scale_color_manual(values=c(scen_cols)[1:3], labels = unique(gsub("(.*)\\.(.*)","\\2", names(init_scen)))) +
  scale_linetype_manual(values=c(1,2), labels=unique(gsub("(.*)\\.(.*)","\\1", names(init_scen)))) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
        legend.position="right", legend.direction="vertical")

initSOCemis

ggsave(paste0(outputdir,"Images/scenario_init.png"), plot= initSOCemis, width = 10, height = 6, dpi = 300, units = "cm")
```

```{r Plot Litter param Scenarios, out.height="6cm", out.width="18cm"}

glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])
glo_SOCStock <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"stock"])

litter_scen <-  c(`LitterPNV-Mean` = "histManagement",
                  `LitterPNV-Grass`= "LitterPNV-PerennialGrasses", 
                  `LitterPNV-AASS` = "LitterPNV-AASS", 
                  `LitterPNV-BRCF` = "LitterPNV-BRCF", 
                  `LitterPNV-MTCF` = "LitterPNV-MTCF", 
                  `LitterPNV-BLEAF`= "LitterPNV-BLEAF",  
                  `LitterPNV-TREF` = "LitterPNV-TREF")

library(ggplot2)
library(scales)
library(gridExtra)
library(cowplot)

theme_set(theme_bw())
legend_size <- 10
scen_cols   <- c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000")[5:1]
input_cols  <- scen_cols[1:4]

limits <- c(-48,-40)
breaks <- c(-40, -42, -44, -46, -48)

litterSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),litter_scen]), 
                                    aes(x = Year, y = Value, group=Data1, color=Data1, size=Data1)) +
  labs( x = "Year", 
        y = expression(paste('Global \u0394SOC in GtC')),
        color="") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous(limits=limits, breaks=breaks) +
  scale_color_manual(values=c("red", scen_cols, "blue"), labels = names(litter_scen)) +
  scale_size_manual(values=c(1,rep(0.5, length(litter_scen)-1)), labels = names(litter_scen)) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
        legend.position = "none")

litterSOCemis

limits <- c(625,705)
breaks <- c(635, 655, 675, 695)

litterSOCstock <- ggplot() + geom_line(data= as.data.frame(glo_SOCStock[,seq(1975,2010,1),litter_scen]/1000), 
                                    aes(x = Year, y = Value, group=Data1, color=Data1, size=Data1)) +
  labs( x = "Year", 
        y = "Global SOC stock\n in GtC",
        color="Litter parameterization",
        size="Litter parameterization") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous(limits=limits, breaks=breaks) +
  scale_color_manual(values=c("red", scen_cols, "blue"), labels = names(litter_scen)) +
  scale_size_manual(values=c(1,rep(0.5, length(litter_scen)-1)), labels = names(litter_scen)) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
        legend.direction="vertical",
        legend.margin=margin(-10,-10,0,0),
        legend.box.margin=margin(-10,-10,0,0)) +
  guides(color=guide_legend(), size = guide_legend())

litterSOCstock

littergrid <- plot_grid(litterSOCemis, litterSOCstock + theme(legend.position = "none"), get_legend(litterSOCstock),
                                  ncol=3, labels=c("(a)","(b)"), axis = "l",
                                  label_fontface = "plain", label_size = 10, align="v")
littergrid

ggsave(paste0(outputdir,"Images/scenario_litter.png"), plot=littergrid, width = 18, height = 6, dpi = 300, units = "cm")

```