### 2022-02-28 SOCbudgetPaper/discussion.Rmd
major update of discussion

# Discussion

We have (1) developed a reduced-complexity model and (2) compiled a spatially explicit time series data set of agricultural management data that helped us to (3) analyze the role of agricultural management in historical cropland SOC dynamics. Our study shows that agricultural management data alter estimates of the SOC stock size and slows down the increase of SOC debt compared to the often used constant management assumptions.

It is important to evaluate the validity of our results as modeling management effects on SOC dynamics at the global scale comes with large uncertainties. The quantitative assessment of the uncertainty of our projections unfortunately exceeds the scope of this article and would likely require a study in itself. The model includes a large number of parameters, and for most of these the uncertainty distributions have not been quantified so far. Moreover, we think that beyond parameter uncertainty, the structural uncertainty from the model design is high. The management data itself is prone to uncertainties as well, as most of it is only indirectly calculated from reported data.

In the following, we give a qualitative assessment of the uncertainties and limitations of this study as well as discuss our three study objectives and results against existing literature.

## Comprehensive historical agricultural management data set

Our spatially explicit time series data set of agricultural management
is based on country-specific FAO production and cropland statistics [@faostat_faostat_2016] as well as 0.5 degree land-use data from LUH2 [@hurtt_harmonization_2020]. Starting from these two source, we compile a harmonized and consistent data set for the major flows within the agricultural system (\@ref(fig:FlowFig)) using rule-based assumptions (e.g. from the IPCC guidelines Vol. 4 [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019]) and only a few auxiliary data sets (e.g. [@porwollik_generating_2018]).

For some of the aspects covered in our data set, for example livestock distribution [@robinson_mapping_2014] or manure production and application [@zhang_global_2017] very well compiled data sets in high resolution exists, that capture real world conditions much better than our estimates. However, they often come with the caveats of either being static in time, demanding large sets of auxiliary data or being inconsistent to each other. Within IAM modeling, not only the SOC model itself, but also the input data that drives the model, has to be computed endogenously and dynamically in time, leaving these data sets, that can not be easily projected into the future unsuitable.

For most of our parameters used in our rule-based calculation no uncertainty estimates exist. This is why, in our view, most of the uncertainty is included in the management data itself, and especially in the residue and manure numbers, as they are only indirectly calculated from production, feed and area data [@faostat_faostat_2016; @weindl_livestock_2017] using large sets of parameter assumptions. The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass. As this is one of the first attempts to compile a comprehensive, global data set on these management data, it is complex to evaluate it against literature. Furthermore, it is still to be evaluated, if our data set covers most of the carbon inputs to the soil, as it not includes cover crops, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar. Unfortunately, data on the quantity of these inputs is very scarce and often does not exist with global coverage.

> (aus den point-by-point replies): In the draft of the International Nitrogen Assessment (will be provided confidentially to the reviewer), the available literature estimates for these flows (in regard to nitrogen) were reviewed. So far no reliable estimates exist, but the existing estimates indicate that these inputs must be by an order of magnitude lower than those of crop residues and manure.

As SOC inputs from above-ground residues had the strongest effect on SOC debt dynamics (see Fig. \@ref(fig:SOCscen)). As pointed out by Keel et al. [-@keel_large_2017] and Smith et al. [-@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study). Keel et al. [-@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources. While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass. Looking on the evaluation of management effects (see Sect. \@ref(sec:ipcccompare)), there is no general indication of overestimating SCFs compared to IPCC default assumptions. In contrast, it is even more likely that we are still missing carbon inputs to the soil from e.g. cover- and inter-cropping practices as SCF are lower than predicted by the IPCC.

## Reduced complexity SOC model

The newly developed reduced-complexity SOC model is based on a Tier 2 modeling approach. This reduces the computational and data demand of the model in comparison to DGVMs. Still, the effects of water supply from spatially heterogenic rainfall and irrigation as well as tillage systems can be accounted for in the computation of SOC decay rates. Similarly, C inputs into the soils reflect differences in productivity in space and over time as well as C losses in the removal rates and processing steps of residues and manure production and application rates to cropland.

The substantial impact of changing management practices in time is indicated by the development of our estimated stock change factors (see Table \@ref(tab:gloSCF)) as well as by the time trend of the SOC debt (see Fig. \@ref(fig:SOCmaps)(a)). Residue management has changed over the last decades, especially with fading out of residue burning practices in several regions and increased general productivity, this shows a clear impact on the SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.

The Tier 2 approach [@ogle_cropland_in_ipcc_2019] used here is explicitly made for agricultural soils, whereas we apply it also to soils under PNV. This is necessary in order to represent SOC losses under land-use change in a dynamics way, which is a main driver of SOC dynamics (cite). Also, the comparison against simulated PNV data with LPJmL4 shows substantial skill in reproducing PNV SOC stocks (Fig. \@ref(fig:SOCLPJmL4)).

Using litterfall estimates from LPJmL4, we have been able to estimate the total SOC stocks of the world, which is dominated by SOC under natural vegetation. However, as the world's SOC stock is highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1; @poggio_soilgrids_2021; @sanderman_soil_2017] in Fig. \@ref(fig:SOCglo), the only conclusion we can draw from this is, that our result is within a plausible range.

To avoid a strong impact of natural land representation and its uncertainties on our results, we focus on SOC changes on cropland. Pristine natural vegetated areas (like permafrost and rain forests) without human land management drop out in our calculation of SOC debt and stock change factors.

Focusing on cropland SOC, some differences between our simulated SOC patterns and reference data show specific shortcomings that need to be considered in the application of our model and in the interpretation of results. In some mountainous regions, we strongly overestimate SOC stocks compared to SoilGrids 2.0 [@poggio_soilgrids_2021]. This could indicate that we are not capable to capture specific processes that would reduce the vegetation's productivity (such as erosion on steep slopes or shallow soils [@borrelli_assessment_2017]) or limited accuracy in the SoilGrids 2.0 data for these regions.

In comparison with aggregated results from the IPCC guidelines, our model shows too strong SOC losses for tropical regions (Table \@ref(tab:SCFglo)). While tropical soils might suffer from low C input rates due to large yield gaps (cite global yield gap atlas?) and high shares of residue removal (MADRAT? FAO for residue burning???), initial SOC stocks under PNV are also comparatively low and thus have little potential for large absolute SOC declines. The revised IPCC factors also indicate that also relative SOC differences are much smaller in the tropics, which cannot be reproduced well by our model. 
With the production-increasing impact of irrigation and fertilization on carbon-poor dryland soils, SOC under cropland can also be higher than under PNV with stock change factors above 1 (see Fig. \@ref(fig:SOCmaps)(d)), but areas are much smaller than areas where the stock change factors are well below unity.

Generally, limiting the analysis to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamic happens in the topsoil. In this regard several aspects are strongly simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Secondly changes to the subsoil due to tillage are neglected. As Powlson et al. [-@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems. No-tillage effects may seem larger than they actually are, if only focusing on the topsoil. SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices. Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.

This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion. However, as pointed out by Doetterl et al. [-@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions [@wang_human-induced_2017]. Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.

> where to put this? With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors (see Table \@ref(tab:gloSCF)) as also indicated by the time trend of the SOC debt. Residue management has changed over the last decades, especially with fading out of residue burning practices in several regions and increased general productivity, this shows a clear impact on the SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.

## SOC debt and SOC drivers

The analysis of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap. There is broad consensus that land conversion to cropland has caused substantial C emissions over the historical period [e.g. @friedlingstein_global_2019]. There is uncertainty with respect to the overall size of these emissions from different methods and reference points and with respect to the contribution of cropland and agricultural management to these emissions. In order to mitigate climate change, there is hope that cropland management can be adapted to sequester atmospheric C in cropland soils [@minasny_soil_2017]. Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], we find that global cropland has acted as a carbon source since 1975. Annual C loss rates of 0.2 per 1000 still have the opposite trend as the promoted 4 per 1000 C sequestration rate target [@minasny_soil_2017]. Dedicated efforts to increase cropland SOC are thus necessary, as management improvements at historical rates are not enough to counteract ongoing SOC degradation on cropland.

According to Sanderman et al. [-@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands. Our estimate of 39.6 GtC in 2010 for cropland debt is thus twice as high as their estimate. However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al. [-@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results.

Furthermore, Sanderman et al. [-@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. Pugh et al. [-@pugh_simulated_2015] considered management effects like tillage and residue returning in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics. Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change). Historical productivity increases were, however, notably larger. Despite large spatial heterogeneity, aggregate yield increase rates are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].

Our study uses a dynamic management data set as driver for SOC dynamics, including historical yield trends and associated increases in residue inputs. [@herzfeld_soc_2021] also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades. Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline. This is consistent with our finding that increasing SOC inputs from above-ground residues had the strongest effect on the slowing-down of the SOC debt increase (Fig. @ref(fig:SOCscen)).

While we find substantially larger effects of productivity gains than the 1% reported by Pugh et al. [-@pugh_simulated_2015], this is not sufficient to compensate SOC losses from moderate global cropland expansion of around 11% between 1974 and 2010.

The effects of agricultural productivity on cropland SOC dynamics can be directly accounted for in our modeling approach, whereas process-based studies [@pugh_simulated_2015; @herzfeld_soc_2021] often lack data on relevant management aspects that drive production increases. Elliott et al. [-@elliott_management_trends_2018] show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, genetic traits, such as kernel number and radiation use efficiency. As such, it will remain challenging for process-based DGVMs to capture the trend of agricultural productivity on cropland SOC dynamics.

Our study emphasizes again that the expansion of croplands is still a major source of CO~2~ emissions --- not only by the removal of vegetation, but also by a slow depletion of C stocks in soils. Our estimates indicate a SOC debt of 39.6 GtC in 2010, and every additional deforestated hectare adds to this debt. Avoided deforestation and other environmental regulation leads to intensification on existing cropland [@humpenoder_large-scale_2018] and our results show that such intensification could lead to increased cropland SOC, if residues are returned to the soil --- amplifying the C sequestration potential of avoided deforestation.

Yet, there is ample potential for further improved SOC management. As shown in Fig. \@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year). Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the soil surface (1.4 GtC). Improved management could include, firstly, a circular flow from the food supply chain back to soils. Waste composting or excreta recycling could represent a major additional N input to cropland soils [@brenzinger_organic_2018]. Secondly, soil carbon sequestration techniques [@smith_soil_2016], deep ploughing [@alcantara_deep_2016] or the transformation of c inputs to more recalcitrant biochar [@woolf_sustainable_2010] may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools. Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs. Finally, other carbon-accumulating practices, such as the cultivation of cover crops [@poeplau_carbon_2015; @porwollik_covercrops_2022] and agroforestry [@lorenz_soil_2014] could increase total C sequestration on cropland. However, data on adoption rates and patterns are missing to help analyzing their impact on historical SOC trends and the future potential of cover crops and agroforestry for increasing SOC stocks has not been explored yet.