### 2022-02-28 ResultNotebooks/create_LUtransition.Rmd
Major update of finalized Paper - Draft 2

---
  title: "LU transitions overview figs"
output:
  md_document:
  variant: markdown_github
---

```{r arrange plot, out.height=3000 , out.width=3300}
library(cowplot)
library(ggplot2)
library(magick)
library(grImport)

path <- "../FlowFigure/"

start     <- "LUC_start.svg"
trans     <- "LUC_trans.svg"
hetero    <- "LUC_hetero.svg"
homo      <- "LUC_homo.svg"
text      <- "LUC_text.svg"

#for(tmp in c(start, trans, hetero, homo, text)) PostScriptTrace(paste0(path, tmp), paste0("Output/", tmp, ".xml"))
for(tmp in c(start, trans, hetero, homo, text)) grConvert::convertPicture(paste0(path, tmp), paste0("Output/", gsub("\\.(.*)", "\\.svg", tmp)))

start     <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", start))))
trans     <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", trans))))
hetero    <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", hetero))))
homo      <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", homo))))
text      <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", text))))

# start     <- pictureGrob(readPicture(paste0("Output/", start, ".xml")))
# trans     <- pictureGrob(readPicture(paste0("Output/", trans, ".xml")))
# hetero    <- pictureGrob(readPicture(paste0("Output/", hetero, ".xml")))
# homo      <- pictureGrob(readPicture(paste0("Output/", homo, ".xml")))
# text      <- pictureGrob(readPicture(paste0("Output/", text, ".xml")))

# gridExtra::grid.arrange(g[1], g[2], g[3], g[4], g[5], nrow=2)

#PostScriptTrace("../FlowFigure/LUC_start.svg")

# # pt <- ggdraw() + draw_image(text)
# 
# #grid_1 <- plot_grid(p1, p2, p3, ncol=3)
# #grid_2 <- plot_grid (pt, p4, ncol=2, rel_widths = c(2, 1))
grid   <- plot_grid(start, trans, hetero, homo, text, nrow=2)
grid
ggsave(paste0(outputdir,"/LUtransition_rotated.pdf"), plot=grid, width = 22, height = 30, units = "cm", bg="white")

```