### 2022-02-28 ResultNotebooks/investigate_SOCvalidate.Rmd

---
title: "SOC budget analysis in comparsion"
output:
  html_document:
    df_print: paged
  pdf_document: default
  bookdown::pdf_document2: default
   keep_md: true
---

```{r load}
library(magclass)
library(madrat)
library(mrcommons)

outputdir <- "Output/Images/"

rev         <- "rev29"

SoilCarbon <- readRDS(paste0("Data/", rev, "histManagement/SoilCarbon.rds"))
Landuse    <- readRDS(paste0("Data/", rev, "histManagement/Landuse.rds"))

```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

cropland        <- Landuse[,,"crop"]
SOCStock_natveg <- dimSums(SoilCarbon[,,c("naturalstate.natveg")], dim=3)
SOCStock_lu     <- dimSums(SoilCarbon[,,c("actualstate")]*Landuse, dim=3)/dimSums(Landuse, dim=3)

SOCStock_natveg[cropland <= 0.001] <- NA
SOCStock_lu[    cropland <= 0.001] <- NA

```

```{r load soc validation stocks}

  # SoilGrids2
  SoilGrids2       <- readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_grid.rds"))
  SoilGrids2_full  <- mbind(setNames(SoilGrids2,     "SOC"),
                            setNames(readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_Q0p95.rds")), "SOC_Q95"),
                            setNames(readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_Q0p05.rds")), "SOC_Q05"))

  #LPJmL4
  LPJmL4_CRU <- toolCoord2Isocell(readRDS(paste0("Data/", rev, "histManagement/LPJmL4_CRU.rds")))
  LPJmL4_CRU <- LPJmL4_CRU[,getYears(Landuse),]

  LPJmL4_CRU[cropland            <= 0.001] <- NA
  SoilGrids2[cropland[,"y2010",] <= 0.001] <- NA
  SoilGrids2_full[cropland[,"y2010",] <= 0.001] <- NA

```

```{r corr LPJmL4 measures}

library(stats)
library(weights)
library(MetricsWeighted)

years <- seq(1975, 2010, 5)

validData <- new.magpie(getCells(SOCStock_natveg), 1, 
                        names = c("LPJmL_CRU", "SOCbudget", "pearson", "meanbias","sdratio", "rmse"), fill = NA)

# for (j in getCells(SOCStock_natveg)) {
#   
#   validData[j,,"pearson"] <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])
#   # validData[j,,"rmse"]    <- MetricsWeighted::rmse(predicted    = SOCStock_natveg[j,years,],
#   #                                                   actual      = LPJmL4_CRU[j,years,],
#   #                                           na.rm = TRUE)
# }

validData[,,"SOCbudget"] <- apply(SOCStock_natveg_raw[,years,], 1, mean)
validData[,,"LPJmL_CRU"] <- apply(LPJmL4_CRU_raw[,years,], 1, mean)
validData[,,"meanbias"]  <- validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"]
validData[,,"sdratio"]   <- apply(SOCStock_natveg_raw[,years,], 1, sd)  / apply(LPJmL4_CRU_raw[,years,], 1, sd)

gloValidData <- new.magpie("GLO", 1, 
                        names = c("socbudget", "lpjml_cru", 
                                  "pearson", "AE", "RMSE", "sdratio", "meanbias"), fill = NA)

gloValidData[,, "socbudget"]     <- dimSums(validData[,,"SOCbudget"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
                                      dim = 1, na.rm = TRUE)

gloValidData[,, "lpjml_cru"]     <- dimSums(validData[,,"LPJmL_CRU"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
                                      dim = 1, na.rm = TRUE)

gloValidData[,, "AE"]      <- dimSums(validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"], 
                                      dim = 1, na.rm = TRUE)

gloValidData[,, "RMSE"]    <- MetricsWeighted::rmse(predicted   = as.vector(validData[,,"SOCbudget"]),
                                                    actual      = as.vector(validData[,,"LPJmL_CRU"]),
                                            na.rm = TRUE, w = as.vector(cropshare[,"y2010",]))

gloValidData[,, "meanbias"] <- stats::weighted.mean(validData[, , "meanbias"],
                                                    w = cropshare[, "y2010", ], na.rm = TRUE)
gloValidData[,, "sdratio"]  <- stats::sd(validData[, , "SOCbudget"], na.rm = TRUE) /
                               stats::sd(validData[, , "LPJmL_CRU"], na.rm = TRUE)
gloValidData[,, "pearson"]  <- weights::wtd.cors(x = validData[,,"SOCbudget"], 
                                                 y = validData[,,"LPJmL_CRU"], 
                                                 weight = cropshare[,"y2010",])

print(gloValidData)

write.magpie(validData,    "validData.rds")
write.magpie(gloValidData, "gloValidData.rds")

#  without time dim
# gewichtet meanbias -> mean value ->map, global -> gewichtet  
# -> RMSE - 
# -> MAE 
# -> Pearson 
# -> sdratio (gewichtet???)

# zeitlich dynamic -> Pearson
# + sdratio -> SI

```

### LPJmL4 - CRU data
```{r LPJmL4_CRU plots, fig.height=6, fig.width=12}

validData_raw <- validData <- validData2    <- read.magpie("validData.rds")
gloValidData <- read.magpie("gloValidData.rds")

brbg_hcl <- colorspace::diverging_hcl(10, h = c(180, 50), c = 80, l = c(20, 95), power = c(0.7, 1.3))

validData[cropshare[,"y2010",]  <= 0.05] <- NA
validData2[cropland[,"y2010",]  <= 0.001] <- NA

luplot::plotmap(validData[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
luplot::plotmap(validData[,,"meanbias"]/ validData[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
luplot::plotmap(validData2[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
luplot::plotmap(validData2[,,"meanbias"]/ validData2[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
luplot::plotmap(validData_raw[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
luplot::plotmap(validData_raw[,,"meanbias"]/ validData_raw[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)

```

```{r quantileSOILGRIDS2, fig.height=6, fig.width=12}

InOrOut <- SOCStock_lu[,"y2010",]
InOrOut[] <- 0
InOrOut[cropshare[,"y2010",] > 0.05] <- 0.5
InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q05"] & 
           SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q95"]] <- 1
InOrOut[SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q05"]] <- 0.3
InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q95"]] <- 0.6
luplot::plotmap(InOrOut)

```

```{r corr soilgrids measures}

library(stats)
library(weights)
library(MetricsWeighted)

years <- 2010

validData <- new.magpie(getCells(SOCStock_natveg), 1, 
                        names = c("LPJmL_CRU", "SOCbudget", "pearson", "meanbias","sdratio", "rmse"), fill = NA)

# for (j in getCells(SOCStock_natveg)) {
#   
#   validData[j,,"pearson"] <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])
#   # validData[j,,"rmse"]    <- MetricsWeighted::rmse(predicted    = SOCStock_natveg[j,years,],
#   #                                                   actual      = LPJmL4_CRU[j,years,],
#   #                                           na.rm = TRUE)
# }

validData[,,"SOCbudget"] <- apply(SOCStock_lu_raw[,years,], 1, mean)
validData[,,"LPJmL_CRU"] <- apply(SoilGrids2_raw[,years,], 1, mean)
validData[,,"meanbias"]  <- validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"]
validData[,,"sdratio"]   <- apply(SOCStock_lu_raw[,years,], 1, sd)  / apply(SoilGrids2_raw[,years,], 1, sd)

gloValidData <- new.magpie("GLO", 1, 
                        names = c("socbudget", "lpjml_cru", 
                                  "pearson", "AE", "RMSE", "sdratio", "meanbias"), fill = NA)

gloValidData[,, "socbudget"]     <- dimSums(validData[,,"SOCbudget"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
                                      dim = 1, na.rm = TRUE)

gloValidData[,, "lpjml_cru"]     <- dimSums(validData[,,"LPJmL_CRU"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
                                      dim = 1, na.rm = TRUE)

gloValidData[,, "AE"]      <- dimSums(validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"], 
                                      dim = 1, na.rm = TRUE)

gloValidData[,, "RMSE"]    <- MetricsWeighted::rmse(predicted   = as.vector(validData[,,"SOCbudget"]),
                                                    actual      = as.vector(validData[,,"LPJmL_CRU"]),
                                            na.rm = TRUE, w = as.vector(cropshare[,"y2010",]))

gloValidData[,, "meanbias"] <- stats::weighted.mean(validData[, , "meanbias"],
                                                    w = cropshare[, "y2010", ], na.rm = TRUE)
gloValidData[,, "sdratio"]  <- stats::sd(validData[, , "SOCbudget"], na.rm = TRUE) /
                               stats::sd(validData[, , "LPJmL_CRU"], na.rm = TRUE)
gloValidData[,, "pearson"]  <- weights::wtd.cors(x = validData[,,"SOCbudget"], 
                                                 y = validData[,,"LPJmL_CRU"], 
                                                 weight = cropshare[,"y2010",])

print(gloValidData)

write.magpie(validData,    "validData_soilgrids.rds")
write.magpie(gloValidData, "gloValidData_soilgrids.rds")

#  without time dim
# gewichtet meanbias -> mean value ->map, global -> gewichtet  
# -> RMSE - 
# -> MAE 
# -> Pearson 
# -> sdratio (gewichtet???)

# zeitlich dynamic -> Pearson
# + sdratio -> SI

```

### LPJmL4 - CRU data
```{r LPJmL4_CRU plots, fig.height=6, fig.width=12}

validData_raw <- validData <- validData2    <- read.magpie("validData_soilgrids.rds")
gloValidData <- read.magpie("gloValidData_soilgrids.rds")

brbg_hcl <- colorspace::diverging_hcl(10, h = c(180, 50), c = 80, l = c(20, 95), power = c(0.7, 1.3))

validData[cropshare[,"y2010",]  <= 0.05] <- NA
validData2[cropland[,"y2010",]  <= 0.001] <- NA

luplot::plotmap(validData[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
luplot::plotmap(validData[,,"meanbias"]/ validData[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
luplot::plotmap(validData2[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
luplot::plotmap(validData2[,,"meanbias"]/ validData2[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
luplot::plotmap(validData_raw[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
luplot::plotmap(validData_raw[,,"meanbias"]/ validData_raw[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)

```