### 2022-02-28 SOCbudgetPaper2/SOCbudget.Rmd

---
title: Management induced changes of soil organic carbon on global croplands 
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Kristine
    surname: Karstens
    affiliation: "1, 3"
    email: kristine.karstens@pik-potsdam.de
    orcid: https://orcid.org/0000-0002-5860-9789
    corresponding: true
  - given_name: Benjamin Leon
    surname: Bodirsky
    affiliation: 1
    email: bodirsky@pik-potsdam.de
    orcid: https://orcid.org/0000-0002-8242-6712
  - given_name: Jan Philipp
    surname: Dietrich
    affiliation: 1 
    orcid: https://orcid.org/0000-0002-4309-6431
  - given_name: Marta
    surname: Dondini
    affiliation: 2
  - given_name: Jens
    surname: Heinke
    affiliation: 1
  - given_name: Matthias
    surname: Kuhnert
    affiliation: 2
  - given_name: Christoph
    surname: Müller
    affiliation: 1 
    orcid: https://orcid.org/0000-0002-9491-3550
  - given_name: Susanne
    surname: Rolinski
    affiliation: 1
  - given_name: Pete
    surname: Smith
    affiliation: 2
    orcid: https://orcid.org/0000-0002-3784-1124
  - given_name: Isabelle
    surname: Weindl
    affiliation: 1
    orcid: https://orcid.org/0000-0002-7651-6930
  - given_name: Hermann
    surname: Lotze-Campen
    affiliation: "1, 3"
  - given_name: Alexander
    surname: Popp
    affiliation: 1
affiliation:
  - code: 1
    address: Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany
  - code: 2
    address: Institute of Biological & Environmental Sciences, University of Aberdeen, Aberdeen, UK
  - code: 3
    address: Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany
abstract: |
  Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth. The first meter of the Earth's soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere. SOC has been depleted by anthropogenic land-cover change and agricultural management. However, the latter has so far not been well represented in global carbon stock assessments. While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale. Here we develop a spatially explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils. We estimate that due to arable farming, soils have lost around 34.6 GtC relative to a counterfactual hypothetical natural state in 1975. Within the period 1975--2010 this SOC debt continued to expand by 5 GtC ($0.14\unit{GtC yr^{-1}}$) to around 39.6 GtC. However, accounting for historical management led to 2.1 GtC less ($0.06\unit{GtC yr^{-1}}$) emissions than under the assumption of constant management. We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique. The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling. 
bibliography: SOCbudget.bib
running:
  title: R Markdown Template for Copernicus
  author: Nüst et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
# copyrightstatement: |
#   The author's copyright for this publication is transferred to Potsdam Institute for Climate Impact Research (PIK).
availability:
  codedata: |
   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \citep{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \citep{karstens_mrsoil_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \citep{bodirsky_mrvalidation_2020} for validation data. All libraries are based on the MADRaT package at github.com/pik-piam/madrat \citep{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing. Model results including C input data are accessable under https://doi.org/10.5281/zenodo.4320663 \citep{karstens_model_2020}. Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.
authorcontribution: |
  KK, BLB and AP designed the study and the model idea. KK wrote the code build on work of BLB, IW. JPD revised and improved the model code. CM, JH and SR provided the LPJmL simulation data. KK wrote the paper with important contributions of BLB and CM. MK, JS, SR and IW provided extensive feedback to outline of the study. All authors discussed the results and commented on the manuscript.

acknowledgements: |
  Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Additional thanks to the rticles contributors for providing a R Markdown template. The authors thank for the data provided by FAOSTAT and LUH2v2. The work of KK was funded by the DFG Priority Program “Climate Engineering: Risks, Challenges, Opportunities?” (SPP 1689) and specifically the CEMICS2 project (grant no. ED78/3-2). The research leading to these results has received funding for BLB from the European Union’s Horizon 2020 research and innovation program under grant agreement no. 776479 (COACCH) and  no. 821010 (CASCADES). The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF). The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no. 774378).

appendix: |

  \section{Figures and tables in appendices}

  \subsection{Methods}

    ```{=latex}
  \begin{figure}[h]
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data. Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
    \label{fig:CLIMzone}
  \end{figure}
  ```

  ```{=latex}
  \begin{sidewaystable*}[htbp]
  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \citet{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions. Allometric coefficients are used as descriped in \citet{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
  }
  \begin{tabular}{lllllllllllll}
  \tophline

           &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
  Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
   \middlehline
  tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
  maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
  trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
  rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
  soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
  rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
  groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
  sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
  oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
  puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
  potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
  cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
  sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
  sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
  others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
  foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
  cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
  \bottomhline
           &                 & \multicolumn{6}{l}{
                                             \begin{minipage}[t]{0.25\columnwidth}
                                                        nr/dm -- nitrogen to dry matter ratio\\
                                                        wm/dm -- wet matter to dry matter ratio\\
                                                        c/dm  -- carbon to dry matter ratio\\
                                                        \end{minipage}} &
                                                                          \multicolumn{5}{l}{
                                                                              \begin{minipage}[t]{0.25\columnwidth}
                                                                                          HI\textsuperscript{area} -- harvest index per area\\
                                                                                          HI\textsuperscript{prod} -- harvest index per production\\
                                                                                          RS -- root:shoot ratio \end{minipage}} 
  \end{tabular}
  (\#tab:c2dm)
  \end{sidewaystable*}
  ```

  ```{=latex}

  \begin{table}[htbp]
    \centering
    \caption{Mapping of FAO crop production items \citep{faostat_faostat_2016} to LUH2 categories \citep{hurtt_harmonization_2020} and to crop groups used in this study.}
    \begin{tabular}{lllll}
    \tophline
        LUH short name & FAO Production Item & Crop group – madrat & Short name – madrat \\ 
    \middlehline    
        c3ann & 122|Sweet potatoes & Tropical roots & cassav\_sp \\ 
        c3ann & 125|Cassava & Tropical roots & cassav\_sp \\ 
        c3ann & 135|Yautia (cocoyam) & Tropical roots & cassav\_sp \\ 
        c3ann & 136|Taro (cocoyam) & Tropical roots & cassav\_sp \\ 
        c3ann & 137|Yams & Tropical roots & cassav\_sp \\ 
        c3ann & "149|Roots and tubers, nes" & Tropical roots & cassav\_sp \\ 
        c3per & 486|Bananas & Tropical roots & cassav\_sp \\ 
        c3per & 489|Plantains & Tropical roots & cassav\_sp \\ 
        c3ann & 328|Seed cotton & Cotton seed & cottn\_pro \\ 
        c3ann & "638|Forage and silage, rye grass" & Forage & foddr \\ 
        c3ann & "639|Forage and silage, grasses nes" & Forage & foddr \\ 
        c3ann & "642|Forage and silage, green oilseeds" & Forage & foddr \\ 
        c3ann & 644|Cabbage for fodder & Forage & foddr \\ 
        c3ann & 645|Pumpkins for Fodder & Forage & foddr \\ 
        c3ann & 646|Turnips for fodder & Forage & foddr \\ 
        c3ann & 647|Beets for fodder & Forage & foddr \\ 
        c3ann & 648|Carrots for fodder & Forage & foddr \\ 
        c3ann & 649|Swedes for fodder & Forage & foddr \\ 
        c3ann & 651|Forage products & Forage & foddr \\ 
        c3ann & 655|Vegetables and roots fodder & Forage & foddr \\ 
        c4ann & "636|Forage and silage, maize" & Forage & foddr \\ 
        c4ann & "637|Forage and silage, sorghum" & Forage & foddr \\ 
        c3nfx & "640|Forage and silage, clover" & Forage & foddr \\ 
        c3nfx & "641|Forage and silage, alfalfa" & Forage & foddr \\ 
        c3nfx & "643|Forage and silage, legumes" & Forage & foddr \\ 
        c3nfx & "242|Groundnuts, with shell" & Groundnuts & groundnut \\ 
        c4ann & 56|Maize & Maize & maiz \\ 
        c3per & "254|Oil, palm fruit" & Oilpalms & oilpalm \\ 
        c3per & 254|Oil palm fruit & Oilpalms & oilpalm \\ 
        c3ann & 358|Cabbages and other brassicas & Fruits Vegetables Nuts & others \\ 
        c3ann & 366|Artichokes & Fruits Vegetables Nuts & others \\ 
        c3ann & 367|Asparagus & Fruits Vegetables Nuts & others \\ 
        c3ann & 372|Lettuce and chicory & Fruits Vegetables Nuts & others \\ 
        c3ann & 373|Spinach & Fruits Vegetables Nuts & others \\ 
        c3ann & 378|Cassava leaves & Fruits Vegetables Nuts & others \\ 
        c3ann & 388|Tomatoes & Fruits Vegetables Nuts & others \\ 
        c3ann & 393|Cauliflowers and broccoli & Fruits Vegetables Nuts & others \\ 
        c3ann & "394|Pumpkins, squash and gourds" & Fruits Vegetables Nuts & others \\ 
        c3ann & 397|Cucumbers and gherkins & Fruits Vegetables Nuts & others \\ 
        c3ann & 399|Eggplants (aubergines) & Fruits Vegetables Nuts & others \\ 
        c3ann & "401|Chillies and peppers, green" & Fruits Vegetables Nuts & others \\ 
        c3ann & "402|Onions, shallots, green" & Fruits Vegetables Nuts & others \\ 
        c3ann & "403|Onions, dry" & Fruits Vegetables Nuts & others \\ 
        c3ann & 406|Garlic & Fruits Vegetables Nuts & others \\ 
        c3ann & "407|Leeks, other alliaceous vegetables" & Fruits Vegetables Nuts & others \\ 
        c3ann & 426|Carrots and turnips & Fruits Vegetables Nuts & others \\ 
        c3ann & 430|Okra & Fruits Vegetables Nuts & others \\ 
        c3ann & 449|Mushrooms and truffles & Fruits Vegetables Nuts & others \\ 
        c3ann & 459|Chicory roots & Fruits Vegetables Nuts & others \\ 
        c3ann & "463|Vegetables, fresh nes" & Fruits Vegetables Nuts & others \\ 
        c3ann & 544|Strawberries & Fruits Vegetables Nuts & others \\ 
        c3ann & 567|Watermelons & Fruits Vegetables Nuts & others \\ 
        c3ann & "568|Melons, other (inccantaloupes)" & Fruits Vegetables Nuts & others \\ 
        c3ann & "689|Chillies and peppers, dry" & Fruits Vegetables Nuts & others \\ 
        c3ann & "711|Anise, badian, fennel, coriander" & Fruits Vegetables Nuts & others \\ 
        c3ann & 720|Ginger & Fruits Vegetables Nuts & others \\ 
        c3ann & "723|Spices, nes" & Fruits Vegetables Nuts & others \\ 
        c3per & "216|Brazil nuts, with shell" & Fruits Vegetables Nuts & others \\ 
        c3per & "217|Cashew nuts, with shell" & Fruits Vegetables Nuts & others \\ 
        c3per & 220|Chestnut & Fruits Vegetables Nuts & others \\ 
        c3per & "221|Almonds, with shell" & Fruits Vegetables Nuts & others \\ 
        c3per & "222|Walnuts, with shell" & Fruits Vegetables Nuts & others \\ 
        c3per & 223|Pistachios & Fruits Vegetables Nuts & others \\ 
        c3per & 224|Kola nuts & Fruits Vegetables Nuts & others \\ 
        c3per & "225|Hazelnuts, with shell" & Fruits Vegetables Nuts & others \\ 
        c3per & 226|Areca nuts & Fruits Vegetables Nuts & others \\ 
        c3per & "234|Nuts, nes" & Fruits Vegetables Nuts & others \\ 
        c3per & 461|Carobs & Fruits Vegetables Nuts & others \\ 
        c3per & 490|Oranges & Fruits Vegetables Nuts & others \\ 
        c3per & "495|Tangerines, mandarins, clementines, satsumas" & Fruits Vegetables Nuts & others \\ 
        c3per & 497|Lemons and limes & Fruits Vegetables Nuts & others \\ 
        c3per & 507|Grapefruit (inc pomelos) & Fruits Vegetables Nuts & others \\ 
        c3per & "512|Fruit, citrus nes" & Fruits Vegetables Nuts & others \\ 
        c3per & 515|Apples & Fruits Vegetables Nuts & others \\ 
        c3per & 521|Pears & Fruits Vegetables Nuts & others \\ 
        c3per & 523|Quinces & Fruits Vegetables Nuts & others \\ 
        c3per & 526|Apricots & Fruits Vegetables Nuts & others \\ 
        c3per & "530|Cherries, sour" & Fruits Vegetables Nuts & others \\ 
        c3per & 531|Cherries & Fruits Vegetables Nuts & others \\ 
        c3per & 534|Peaches and nectarines & Fruits Vegetables Nuts & others \\ 
        c3per & 536|Plums and sloes & Fruits Vegetables Nuts & others \\ 
        c3per & "541|Fruit, stone nes" & Fruits Vegetables Nuts & others \\ 
        c3per & "542|Fruit, pome nes" & Fruits Vegetables Nuts & others \\ 
        c3per & 547|Raspberries & Fruits Vegetables Nuts & others \\ 
        c3per & 549|Gooseberries & Fruits Vegetables Nuts & others \\ 
        c3per & 550|Currants & Fruits Vegetables Nuts & others \\ 
        c3per & 552|Blueberries & Fruits Vegetables Nuts & others \\ 
        c3per & 554|Cranberries & Fruits Vegetables Nuts & others \\ 
        c3per & 558|Berries nes & Fruits Vegetables Nuts & others \\ 
        c3per & 560|Grapes & Fruits Vegetables Nuts & others \\ 
        c3per & 569|Figs & Fruits Vegetables Nuts & others \\ 
        c3per & "571|Mangoes, mangosteens, guavas" & Fruits Vegetables Nuts & others \\ 
        c3per & 572|Avocados & Fruits Vegetables Nuts & others \\ 
        c3per & 574|Pineapples & Fruits Vegetables Nuts & others \\ 
        c3per & 577|Dates & Fruits Vegetables Nuts & others \\ 
        c3per & 587|Persimmons & Fruits Vegetables Nuts & others \\ 
        c3per & 591|Cashewapple & Fruits Vegetables Nuts & others \\ 
        c3per & 592|Kiwi fruit & Fruits Vegetables Nuts & others \\ 
        c3per & 600|Papayas & Fruits Vegetables Nuts & others \\ 
        c3per & "603|Fruit, tropical fresh nes" & Fruits Vegetables Nuts & others \\ 
        c3per & "619|Fruit, fresh nes" & Fruits Vegetables Nuts & others \\ 
        c3per & "656|Coffee, green" & Fruits Vegetables Nuts & others \\ 
        c3per & "661|Cocoa, beans" & Fruits Vegetables Nuts & others \\ 
        c3per & 667|Tea & Fruits Vegetables Nuts & others \\ 
        c3per & 671|Mate & Fruits Vegetables Nuts & others \\ 
        c3per & 687|Pepper (piper spp) & Fruits Vegetables Nuts & others \\ 
        c3per & 692|Vanilla & Fruits Vegetables Nuts & others \\ 
        c3per & 693|Cinnamon (canella) & Fruits Vegetables Nuts & others \\ 
        c3per & 698|Cloves & Fruits Vegetables Nuts & others \\ 
        c3per & "702|Nutmeg, mace and cardamoms" & Fruits Vegetables Nuts & others \\ 
        c4ann & "446|Maize, green" & Fruits Vegetables Nuts & others \\ 
        c3nfx & "414|Beans, green" & Fruits Vegetables Nuts & others \\ 
        c3nfx & "417|Peas, green" & Fruits Vegetables Nuts & others \\ 
        c3nfx & "420|Vegetables, leguminous nes" & Fruits Vegetables Nuts & others \\ 
        c3nfx & 423|String beans & Fruits Vegetables Nuts & others \\ 
        c3ann & 677|Hops & Fruits Vegetables Nuts & others \\ 
        c3ann & 748|Peppermint & Fruits Vegetables Nuts & others \\ 
        c3ann & 116|Potatoes & Potatoes & potato \\ 
        c3nfx & "176|Beans, dry" & Pulses & puls\_pro \\ 
        c3nfx & "181|Broad beans, horse beans, dry" & Pulses & puls\_pro \\ 
        c3nfx & "187|Peas, dry" & Pulses & puls\_pro \\ 
        c3nfx & 191|Chick peas & Pulses & puls\_pro \\ 
        c3nfx & "195|Cow peas, dry" & Pulses & puls\_pro \\ 
        c3nfx & 197|Pigeon peas & Pulses & puls\_pro \\ 
        c3nfx & 201|Lentils & Pulses & puls\_pro \\ 
        c3nfx & 203|Bambara beans & Pulses & puls\_pro \\ 
        c3nfx & 205|Vetches & Pulses & puls\_pro \\ 
        c3nfx & 210|Lupins & Pulses & puls\_pro \\ 
        c3nfx & "211|Pulses, nes" & Pulses & puls\_pro \\ 
        c3ann & 265|Castor oil seed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 270|Rapeseed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 280|Safflower seed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 289|Sesame seed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 292|Mustard seed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 296|Poppy seed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 299|Melonseed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 333|Linseed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & 336|Hempseed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 249|Coconuts & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 260|Olives & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 263|Karite nuts (sheanuts) & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 275|Tung nuts & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 277|Jojoba seed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 305|Tallowtree seed & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 310|Kapok fruit & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 311|Kapokseed in shell & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3per & 339|Oilseeds nes & Other oil crops (incl rapeseed) & rapeseed \\ 
        c3ann & "27|Rice, paddy" & Rice & rice\_pro \\ 
        c3nfx & 236|Soybeans & Soybean & soybean \\ 
        c3ann & 157|Sugar beet & Sugar beet & sugr\_beet \\ 
        c4per & 156|Sugar cane & Sugar cane & sugr\_cane \\ 
        c3ann & 267|Sunflower seed & Sunflower & sunflower \\ 
        c3ann & 15|Wheat & Temperate cereals & tece \\ 
        c3ann & 44|Barley & Temperate cereals & tece \\ 
        c3ann & 71|Rye & Temperate cereals & tece \\ 
        c3ann & 75|Oats & Temperate cereals & tece \\ 
        c3ann & 89|Buckwheat & Temperate cereals & tece \\ 
        c3ann & 92|Quinoa & Temperate cereals & tece \\ 
        c3ann & 94|Fonio & Temperate cereals & tece \\ 
        c3ann & 97|Triticale & Temperate cereals & tece \\ 
        c3ann & 101|Canary seed & Temperate cereals & tece \\ 
        c3ann & "103|Grain, mixed" & Temperate cereals & tece \\ 
        c3ann & "108|Cereals, nes" & Temperate cereals & tece \\ 
        c4ann & 68|Popcorn & Temperate cereals & tece \\ 
        c4ann & 79|Millet & Tropical cereals & trce \\ 
        c4ann & 83|Sorghum & Tropical cereals & trce \\ 
        \bottomhline
    \end{tabular}
    (\#tab:fao2luh2mag)
  \end{table}
  ```

  <!-- closs in AWMS - Table {#append:TableclossAWMS} -->

  \subsection{Discussion}

    ```{=latex}
  \begin{figure}[H]
    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
    \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010. PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model. We use the climate zone specification of the IPCC \citep{eggleston_ipcc_2006}.}
    \label{fig:SOCreg}
  \end{figure}
  ```

output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
    md_extensions: +raw_attribute
  md_document:
    variant: markdown_github
  bookdown::word_document2: default
  rticles::copernicus_article: default
  bookdown::pdf_document2: default
---
\newpage

```{r child = 'introduction.Rmd'}
```
\newpage

```{r child = 'methods.Rmd'}
```
\newpage

```{r child = 'results.Rmd'}
```
\newpage

```{r child = 'discussion.Rmd'}
```
\newpage

```{r child = 'conclusion.Rmd'}
```