### 2022-02-28 SOCbudgetPaper2/conclusion.Rmd

\conclusions

We have compiled a spatially explicit and time-variant data set on agricultural management aspects relevant for cropland SOC dynamics. We have also developed a reduced-complexity SOC model that is able to be applied in optimization-based IAM frameworks, for which detailed process-based models are computationally too expensive. Making use of these data and model, we are able to estimate spatially explicit SOC stocks, SOC debts, and stock change factors considering agricultural management. It is --- to our knowledge --- the first study that analyzes the role of time-variant and spatially explicit historical agricultural management for global SOC dynamics.

Our results demonstrate that historical changes in agricultural management have shaped the SOC debt on cropland. It is thus necessary to explicitly consider agricultural management in a dynamic manner in global carbon assessments and models, especially when exploring climate mitigation pathways with so-called land-based solutions [e.g. @popp_land-use_2016]. That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data. Our open-source model [@karstens_mrsoil_2020], published data-set [@karstens_model_2020] and the flexible data processing with the MADRaT package [@dietrich_madrat_2020] constitute a starting point for building comprehensive data sets on agricultural management aspects.

With the reduced-complexity SOC model we are able to account for agricultural management effects on cropland SOC dynamics within optimization-based IAM frameworks. Reduced input data requirements such as accounting for changes in productivity rather than reproducing the processes that lead to such changes in productivity [@elliott_management_trends_2018] will help to explore the role of agricultural management in sustainable development pathway analyses [@sorgel_sustainable_2021].
However, we clearly see that increases in agricultural productivity are not sufficient to create positive net SOC sequestration in cropland soils. More management options that explicitly target the sequestration of C in cropland soils need to be considered. Our open-source model can be expanded to account for additional management options for carbon farming, such as cover crops, agroforestry, or biochar applications.