### 2022-07-20 ResultNotebooks/create_ZenodoOutput.Rmd
update of output scripts and figure

---
  title: "SOC budget aóutput creator"
output:
  html_document:
  df_print: paged
pdf_document: default
bookdown::pdf_document2: default
keep_md: true
---

Load all folder and create NetCDF files:

```{r load files}

rev        <- "Data/rev30"
scenarios   <- c("histManagement", "constManagement-1975", "constManure-1975", "constResidues-1975", "constTillage-mixedtill")

for (s in scenarios[1]) {

   path  <- paste0(rev, s)
   files <- list.files(path, pattern = "*rds")
   # dir.create(s)

   for (f in files) {

     tmp <- readRDS(paste0(path, "/", f))
     f   <- gsub(".rds", "", f)
     items <- getItems(tmp)

     if (length(unlist(items[1])) == 59199) output <- ".nc" else output <- ".csv"
     print(paste(s, f))
     print(unlist(items[2])[c(1, length(unlist(items[2])))])
     print(unname(unlist(items[3])))

     # write.magpie(tmp, file_name = paste0(f, output), file_folder = paste0(outputdir, s))
   }

   # tar(paste0(s, ".tgz"), files = s, compression = "gzip")

}
```