### 2022-07-20 ResultNotebooks/results_SOCattribution.Rmd

---
  title: "Global SOC scenarios"
output:
  md_document:
  variant: markdown_github
html_document:  
  pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrsoil)
library(magpiesets)

rev         <- "rev30"
scenarios   <- c("histManagement", "constManagement-1975")
filesNeeded <- c("SoilCarbon.rds", "Landuse.rds")
outputdir   <- "Output/"
```

```{r Load data and calc globals}

SOCscen <- NULL

for (scen in scenarios) {
  print(scen)
  SoilCarbon <- read.magpie(
                  paste0("Data/", paste0(rev, scen), "/SoilCarbon.rds"))
  Landuse    <- read.magpie(
                  paste0("Data/", paste0(rev, scen), "/Landuse.rds"))
  SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
  SOCStock   <- toolConditionalReplace(dimSums(dimSums(
    SoilCarbon[, , c("actualstate", "naturalstate")], dim = 3.3) *
      Landuse, dim = 3.2), "is.nan()", 0)
  SOCscen    <- mbind(SOCscen, 
                      add_dimension(SOCStock,
                                    dim = 3.1, nm = scen, add = "scenario"))
}

SOCDebt       <- collapseNames(SOCscen[, , "actualstate"] -
                                 SOCscen[, , "naturalstate"])
SOCDebtChange <- collapseDim(SOCDebt[, c(1975:2010), ] -
                               SOCDebt[, "y1975", ])

landMask   <- collapseDim(Landuse[, 1, 1])
landMask[] <- NA
for (y in c(1975:2010)) landMask[Landuse[, y, "crop"] > 0.001] <- 1

LUCattr <- toolConditionalReplace(
             SOCDebtChange[, "y2010", "constManagement-1975"],
             "is.na()", 0) * landMask

MANattr <- toolConditionalReplace(
             (SOCDebtChange[, "y2010", "histManagement"] -
              SOCDebtChange[, "y2010", "constManagement-1975"]),
             "is.na()", 0) * landMask

saveRDS(
  mbind(
    add_dimension(collapseNames(SOCDebt[, "y2010", "histManagement"]) *
                    landMask, dim = 3.1, add = "type", nm = "SOCDebt"),
    add_dimension(collapseNames(SOCDebtChange[, "y2010", "histManagement"]) *
                    landMask, dim = 3.1, add = "type", nm = "SOCDebtChange"),
    add_dimension(collapseNames(LUCattr),
                  dim = 3.1, add = "type", nm = "LUCdiff"),
    add_dimension(collapseNames(MANattr),
                  dim = 3.1, add = "type", nm = "MANdiff")),
  paste0(outputdir, "SOCDebtSpatial.rds"))

# LUCattr <- toolConditionalReplace(
#              SOCDebtChange[,"y2010", "constManagement-1975"] /
#              SOCDebtChange[,"y2010","histManagement"],
#              "is.na()", 0) * landMask
#
# MANattr <- toolConditionalReplace(
#              (SOCDebtChange[,"y2010","histManagement"] -
#               SOCDebtChange[,"y2010", "constManagement-1975"]) /
#               SOCDebtChange[,"y2010","histManagement"],
#              "is.na()", 0) * landMask
#
# signLUC   <- landMask
# signLUC[] <- 1
# signLUC[SOCDebtChange[,"y2010","constManagement-1975"] < 0] <- (-1)
#
# signMAN   <- landMask
# signMAN[] <- 1
# signMAN[(SOCDebtChange[,"y2010","histManagement"] -
#            SOCDebtChange[,"y2010","constManagement-1975"]) < 0] <- (-1)
#
# signLUCattr <- collapseNames(abs(LUCattr) * signLUC)
# signMANattr <- collapseNames(abs(MANattr) * signMAN)
#
# saveRDS(mbind(add_dimension(signLUCattr, dim = 3.1,
#                             add = "type", nm = "LUCattr"),
#               add_dimension(signMANattr, dim = 3.1,
#                             add = "type", nm = "MANattr")),
#         paste0(outputdir, "SOCattribution.rds"))
```

```{r grid plot, out.height=800 , out.width=1200}

library(reshape2)

SOCDebtSpatial <- readRDS(paste0(outputdir, "SOCDebtSpatial.rds"))
magpie_coord   <- toolGetMapping("magpie_coord.rda",
                                type = "cell", where = "mappingfolder")

SOCDebt2plot       <- 
  merge(as.data.frame(SOCDebtSpatial[, , "SOCDebt"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
SOCDebtChange2plot <- 
  merge(as.data.frame(SOCDebtSpatial[, , "SOCDebtChange"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
LUCattri2plot      <- 
  merge(as.data.frame(SOCDebtSpatial[, , "LUCdiff"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
MANattri2plot      <- 
  merge(as.data.frame(SOCDebtSpatial[, , "MANdiff"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")

library(ggplot2)
theme_set(theme_bw())
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(sp)
library(rworldmap)
library(oce)
library(scales)
library(cowplot)

legend_size <- 10

map.world <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)
coastline.world <- data.frame(longitude = coastlineWorld[["longitude"]],
                              latitude = coastlineWorld[["latitude"]])

SOCplotmap   <- function(data, name, cols, limits,
                         values = NULL, breaks, labels, legend_width) {

out <- ggplot(data = world) + geom_sf() +
              geom_tile(data = subset(data, Year == "2010", 
                                      c("lon", "lat", "Value")),
                        aes(x = lon, y = lat, fill = Value)) +
        scale_fill_stepsn(limits   = limits,
                      colours  = cols,
                      name     = name,
                      values   = values,
                      breaks   = breaks,
                      labels   = labels,
                      na.value = "white",
                      guide    = guide_legend(
                        title.theme    = element_text(size = legend_size),
                        direction      = "horizontal",
                        title.position = "top",
                        # I shift the labels around, the should be placed
                        # exactly at the right end of each legend key
                        title.hjust    = 0.5,
                        label.hjust    = 1,
                        nrow           = 1,
                        byrow          = FALSE,
                        label.position = "bottom",
                        keyheight      = 0.5,
                        keywidth       = legend_width
                      )) +
    theme(legend.position = "bottom",
          axis.title.x     = element_blank(),
          axis.title.y     = element_blank(),
          legend.text      = element_text(size = legend_size),
          rect             = element_blank(),
          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7, 0.7)),
          plot.margin      = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

    return(out)
}

cols     <- colorRampPalette(brewer.pal(11, "RdYlBu"))(11)[c(1:6, 7, 9, 11)]
lim_debt <- 10
val_debt <- c(-10, -6, -4, -2, -1, 0, 1, 10)
debt          <- 
  SOCplotmap(data     = SOCDebt2plot,
             limits   = c(-lim_debt, lim_debt),
             cols     = cols,
             name     = expression("\u0394SOC in 2010 in million tons of carbon"),
             values   = (val_debt + lim_debt) / (2 * lim_debt),
             breaks   = val_debt[-1],
             labels   =  c("< -6", "    -4", "    -2", "    -1 ", "   0", "  1", "> 1"),
             legend_width = 2.3)

cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(10)
lim_abs       <- 1.5
val_abs       <- seq(-lim_abs, lim_abs, length.out = 11)
debttrend_abs <-
  SOCplotmap(data     = SOCDebtChange2plot,
             limits   = c(-lim_abs, lim_abs),
             cols     = cols,
             name     = expression("\u0394SOC change from 1975 to 2010 in million tons of carbon"),
             values   = (val_abs + lim_abs) / (2 * lim_abs),
             breaks   = val_abs[-1],
             labels   = c(paste0("< ", val_abs[2]),
                          as.character(val_abs[c(-1, -2, -11)]),
                          paste0("> ", val_abs[10])),
             legend_width = 1.5)

debttrend_luc <- 
  SOCplotmap(data     = LUCattri2plot,
             limits   = c(-lim_abs, lim_abs),
             cols     = cols,
             name     = expression("\u0394SOC change from 1975 to 2010 in million tons of carbon"),
             values   = (val_abs + lim_abs) / (2 * lim_abs),
             breaks   = val_abs[-1],
             labels   = c(paste0("< ", val_abs[2]),
                          as.character(val_abs[c(-1, -2, -11)]),
                          paste0("> ", val_abs[10])),
             legend_width = 1.5)

debttrend_man <- 
  SOCplotmap(data     = MANattri2plot,
             limits   = c(-lim_abs, lim_abs),
             cols     = cols,
             name     = expression("\u0394SOC change from 1975 to 2010 in million tons of carbon"),
             values   = (val_abs + lim_abs) / (2 * lim_abs),
             breaks   = val_abs[-1],
             labels   = c(paste0("< ", val_abs[2]),
                          as.character(val_abs[c(-1, -2, -11)]),
                          paste0("> ", val_abs[10])),
             legend_width = 1.5)

debt
debttrend_abs
debttrend_luc
debttrend_man

library(gridExtra)
library(cowplot)

grid  <- plot_grid(debt,
                   debttrend_abs + guides(fill = "none"),
                   debttrend_luc + guides(fill = "none"),
                   debttrend_man, ncol = 1,
                   labels = c("(a)  - \u0394SOC ",
                              "(b1) - total \u0394SOC change",
                              "(b2) - \u0394SOC change due to LUC",
                              "(b3) - \u0394SOC change due to MAN"),
                   label_fontface = "plain", label_size = 10,
                   align = "v",  rel_heights = c(1.40, 1, 1, 1.40),
                   label_x = 0.01, hjust = 0, vjust = 1.1
                   )

grid

ggsave(paste0(outputdir, "Images/SpatialDebtFigure.png"), plot = grid,
       width = 10, height = 26, dpi = 300, units = "cm")
```