### 2022-07-20 ResultNotebooks/results_SOCcompare.Rmd

---
title: "SOC budget analysis in comparsion to other global SOC estimates"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r load}
library(magclass)
library(madrat)

rev        <- "rev30"
outputdir  <- "Output/"

SoilCarbon <- read.magpie(paste0("Data/", rev, "histManagement/SoilCarbon.rds"))
Landuse    <- read.magpie(paste0("Data/", rev, "histManagement/Landuse.rds"))
```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

SoilCarbon[, , c("naturalstate.crop")] <- SoilCarbon[, , c("naturalstate.natveg")]
SOCStock     <- dimSums(dimSums(SoilCarbon[, , c("naturalstate", "actualstate")],
                                dim = 3.3) * Landuse, dim = 3.2)
cell2reg     <- read.csv("Data/additional/ClimateZoneMapping.csv")
reg_SOCStock <- toolAggregate(SOCStock, rel = cell2reg, from = "cell", to = "IPCC")
```

```{r table glo soc stocks}

WISE         <- readRDS(paste0("Data/",
                                 rev, "histManagement/WISE.rds"))
  GSOC         <- readRDS(paste0("Data/",
                                 rev, "histManagement/GSOC.rds"))
  SoilGrids    <- readRDS(paste0("Data/",
                                 rev, "histManagement/SoilGrids.rds"))
  SoilGrids2   <- readRDS(paste0("Data/",
                                 rev, "histManagement/SoilGrids2.rds"))
  LPJmL4       <- readRDS(paste0("Data/",
                                 rev, "histManagement/LPJmL4.rds"))
  LPJmL4_CRU   <- readRDS(paste0("Data/",
                                 rev, "histManagement/LPJmL4_CRU.rds"))
  SOCDebtPaper <- readRDS(paste0("Data/",
                                 rev, "histManagement/SOCDebtPaperC.rds"))

library(knitr)
library(stats)
library(reshape2)

SOCStock <- reg_SOCStock
reg_SOC  <-
  mbind(setNames(SOCStock[, "y2010", "actualstate"], "This Study"),
        setNames(SOCStock[, "y2010", "naturalstate"], "This Study (PNV)"),
        setNames(LPJmL4[, "y2010", ], "LPJmL4"),
        setYears(setNames(WISE[, 1, ], "WISE"), "y2010"),
        setYears(setNames(SoilGrids[, 1, ], "SoilGrids"), "y2010"),
        setYears(setNames(GSOC[, 1, ], "GSOC"), "y2010"),
        setYears(setNames(SoilGrids2[, 1, ], "SoilGrids 2.0"), "y2010"),
        setYears(setNames(SOCDebtPaper[, "y2010", ], "SOCDebtPaper"), "y2010"))
glo_SOC <- mbind(setCells(dimSums(reg_SOC, dim = 1), "GLO"), reg_SOC)

# print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))

comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(glo_SOC / 1000)),
                           Data1 ~ Region, value.var = "Value",
                            row.names = levels(as.data.frame(glo_SOC)$Data1)))

comparison_table$Year <- "2010"

saveRDS(object = comparison_table, paste0(outputdir, "gloSOC_comparison.rds"))
```

```{r point plot reg stock, out.width=800, out.height=600}
library(ggplot2)
library(knitr)
library(stats)
library(reshape2)

table    <- readRDS("Output/gloSOC_comparison.rds")
leg_name <- ""

names(table) <- c("Data1", "Boreal Dry", "Boreal Moist", "C-Temp Dry",
                  "C-Temp Moist", "GLO", "Polar Dry", "Polar Moist",
                  "Tropical Dry", "Tropical Moist", "Tropical Montane",
                  "Tropical Wet", "W-Temp Dry", "W-Temp Moist", "Year")
table        <- table[, c(1, 15, 6, 7, 8, 2, 3, 4, 5, 13, 14, 9, 10, 11, 12)]

cols   <- c("black",   "#DF536B", "#61D04F",
           "#2297E6", "#28E2E5", "#CD0BBC",
           "#F5C710", "gray62")[c(2, 7, 1, 3, 4, 5, 6, 8)]

glob <- ggplot(data = melt(subset(table, select = c(Data1, GLO, Year))),
               aes(x = Year, y = value, group = Data1)) +
  geom_point(aes(colour = Data1, size = Data1, shape = Data1)) +
  scale_size_manual(values = c(4,  2, 2, 2, 2, 2, 2,  2)) +
  scale_shape_manual(values = c(19, 16, 6, 2, 5, 0, 8, 10)) +
  scale_color_manual(values = cols) +
  labs(color  = leg_name, size = leg_name, shape = leg_name) +
  xlab("") + ylab("SOC (0-30cm) in GtC") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x  = element_blank(),
        legend.position = "right")

comp <- ggplot(data = melt(subset(table, select = -GLO)),
               aes(x = Year, y = value, group = Data1)) +
  geom_point(aes(colour = Data1, size = Data1, shape = Data1)) +
  facet_wrap(~variable, scales = "free", drop = TRUE,
             labeller = label_wrap_gen(width = 18)) +
  scale_size_manual(values = c(4,  2, 2, 2, 2, 2, 2,  2)) +
  scale_shape_manual(values = c(19, 16, 6, 2, 5, 0, 8, 10)) +
  scale_color_manual(values = cols) +
  labs(color  = leg_name, size = leg_name, shape = leg_name) +
  xlab("") + ylab("SOC (0-30cm) in GtC") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x  = element_blank(),
        legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 3, byrow = TRUE))

glob
comp

ggsave(paste0(outputdir, "/Images/glo_comparisonfigure.png"),
       plot = glob, height =  7, width =  7, units = "cm", dpi = 300)
ggsave(paste0(outputdir, "/Images/reg_comparisonfigure.png"),
       plot = comp, height = 15, width = 14, units = "cm", dpi = 300)
```

```{r point plot glo stock, fig.width=3, fig.height=4}
library(ggplot2)

glob
ggsave(paste0(outputdir, "/glo_comparisonfigure.png"), plot = glob)
```
`