### 2022-07-20 ResultNotebooks/results_SOCmaps.Rmd

---
title: "Global SOC maps and emissions"
output:
  md_document:
    variant: markdown_github
---

```{r load}
library(magclass)
library(madrat)
library(magpiesets)

rev         <- "rev30"
filesNeeded <- c("SoilCarbon.rds", "Landuse.rds")
scenarios   <- c("histManagement")
outputdir   <- "Output/Images"

SoilCarbon <- read.magpie(paste0("Data/", rev, scenarios[1], "/SoilCarbon.rds"))
Landuse    <- read.magpie(paste0("Data/", rev, scenarios[1], "/Landuse.rds"))
```

```{r Carbon Emissions (glo)}

library(reshape2)

years <- seq(1975, 2010, 1)

SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
SOCStock        <- dimSums(
                     dimSums(
                       SoilCarbon[, years, ][, , "state", pmatch = TRUE],
                       dim = 3.3) *
                     Landuse[, years, ],
                     dim = 3.2)
glo_SOCStock    <- dimSums(SOCStock, dim = 1)
glo_SOCEmission <- (glo_SOCStock[, , "naturalstate"] -
                      glo_SOCStock[, , "actualstate"]) / 1000

cropland_dens   <- dimSums(SoilCarbon[, , "actualstate"][, , "crop"], dim = 3)
cropland_dens[Landuse[, , "crop"] <= 0.001] <- NA

cropland_share <- dimSums(SoilCarbon[, , "actualstate"][, , "crop"],
                          dim = 3) /
                 (dimSums(SoilCarbon[, , "naturalstate"][, , "natveg"],
                          dim = 3) + 1e-8)
cropland_share[Landuse[, , "crop"] <= 0.001] <- NA

cropland_increase <- dimSums(SoilCarbon[, , "actualstate"][, , "crop"],
                             dim = 3) -
                     dimSums(SoilCarbon[, , "naturalstate"][, , "natveg"],
                             dim = 3)
cropland_increase[Landuse[, , "crop"] <= 0.001] <- NA

magpie_coord  <- toolGetMapping("magpie_coord.rda", type = "cell")
share2plot    <- merge(as.data.frame(cropland_share),
                       data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
increase2plot <- merge(as.data.frame(cropland_increase),
                       data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
density2plot  <- merge(as.data.frame(cropland_dens),
                       data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
```

```{r grid plot, out.height=800 , out.width=1200}

library(ggplot2)
theme_set(theme_bw())
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(sp)
library(rworldmap)
library(oce)
library(scales)

legend_size <- 10

map.world <- map_data("world")
world     <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)
coastline.world <- data.frame(longitude = coastlineWorld[["longitude"]],
                              latitude  = coastlineWorld[["latitude"]])

squish_trans <- function(from, to, factor) {

  trans <- function(x) {
    if (any(is.na(x))) return(x)
    # get indices for the relevant regions
    isq <- x > from & x < to
    ito <- x >= to
    # apply transformation
    x[isq] <- from + (x[isq] - from) / factor
    x[ito] <- from + (to - from) / factor + (x[ito] - to)
    return(x)
  }

  inv <- function(x) {
    if (any(is.na(x))) return(x)
    # get indices for the relevant regions
    isq <- x > from & x < from + (to - from) / factor
    ito <- x >= from + (to - from) / factor
    # apply transformation
    x[isq] <- from + (x[isq] - from) * factor
    x[ito] <- to + (x[ito] - (from + (to - from) / factor))
    return(x)
  }
  # return the transformation
  return(trans_new("squished", trans, inv))
}

SOCemis <- ggplot() +
  geom_line(data = as.data.frame(-glo_SOCEmission)[, c(3, 6)],
            aes(x = Year, y = Value, group = 1)) +
  labs(x = "Year",
        y = expression(paste("Global \u0394SOC in GtC"))) +
  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
  scale_y_continuous(trans = squish_trans(-34, 0, 10),
                     limits = c(-40, 0),
                     breaks = c(0, -34, -35, -36, -37, -38, -39, -40),
                     minor_breaks = seq(-1, -29, -2)) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size),
        axis.title.x = element_text(size = legend_size,
                                    margin = margin(t = 10,
                                                    r = 0,
                                                    b = 10,
                                                    l = 0,
                                                   unit = "pt")),
        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8,
                                   size = legend_size)
        ) +
  annotate(geom = "text", label = "SOC debt",
           x = factor(1995), y = -20, size = 3) +
  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -36,
           arrow = arrow(end    = "both",
                         type   = "closed",
                         length = unit(0.1, "inches")))

SOCplotmap   <- function(data, name, cols, limits,
                         values = NULL, breaks, labels) {

  out <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(data, Year == "2010", c("lon", "lat", "Value")),
              aes(x = lon, y = lat, fill = Value)) +
    scale_fill_stepsn(limits   = limits,
                      colours  = cols,
                      name     = name,
                      values   = values,
                      breaks   = breaks,
                      labels   = labels,
                      na.value = "white",
                      guide    = guide_legend(
                        title.theme    = element_text(size = legend_size),
                        direction      = "horizontal",
                        title.position = "top",
                        # I shift the labels around, the should be placed
                        # exactly at the right end of each legend key
                        title.hjust    = 0.5,
                        label.hjust    = 1,
                        nrow           = 1,
                        byrow          = FALSE,
                        label.position = "bottom",
                        keyheight      = 0.5,
                        keywidth       = 2
                      )) +
    theme(legend.position = "bottom",
          axis.title.x     = element_blank(),
          axis.title.y     = element_blank(),
          legend.text      = element_text(size = legend_size),
          rect             = element_blank(),
          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7, 0.7)),
          plot.margin      = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

    return(out)
}

cols        <- colorRampPalette(brewer.pal(11, "RdYlBu"))(11)[c(1:6, 7, 9, 11)]
SOCshare    <-
  SOCplotmap(
    data     = share2plot,
    limits   = c(0, 2),
    cols     = cols,
    name     = expression("cropland SOC stock change factor (unitless)"),
    values   = c(0, 0.2, 0.4, 0.6, 0.8, 1.0, 1.2, 2) / 2,
    breaks   = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.2, 2),
    labels   = c("< 0.2", "0.4", "0.6", "0.8", "1.0", "1.2", "> 1.2"))

SOCincrease <-
  SOCplotmap(
    data     = increase2plot,
    limits   = c(-100, 100),
    cols     = cols,
    name     = expression(
                 paste("cropland \u0394SOC in tons of carbon per hectare")),
    values   = (c(-100, -40, -30, -20, -10, 0, 10, 100) + 100) / 200,
    breaks   =  c(-40, -30, -20, -10, 0, 10, 100),
    labels   =  c("< -40", "        -30", "        -20",
                  "      -10 ", "     0", "    10", "> 10"))

cols <-
  colorRampPalette(
    c("white",
      brewer.pal(9, "YlGn")[c(1:2)],
      "yellow",
      brewer.pal(9, "YlGn")[c(4:7)],
      brewer.pal(9, "YlGnBu")[c(6:9)],
      "black"))(21)[c(1, 5, 7, 10, 13, 16, 19) + 1]

SOCdensity  <-
  SOCplotmap(
    data     = density2plot,
    limits   = c(0, 150),
    cols     = cols,
    name     = expression("cropland SOC in tons of carbon per hectare"),
    values   = c(0, 5, 10, 20, 40, 70, 100, 150) / 150,
    breaks   = c(5, 10, 20, 40, 70, 100, 150),
    labels   = c("0     5", "  10", "  20", "  40", "  70", "  100", " >150"))

library(gridExtra)
library(cowplot)

SOCemis
SOCincrease
SOCdensity
SOCshare

grid <- plot_grid(SOCemis, SOCdensity, SOCincrease, SOCshare,
                  ncol = 2, labels = c("(a)", "(b)", "(c)", "(d)"),
                  label_fontface = "plain", label_size = 10)

grid
ggsave(paste0(outputdir, "/4panelfigure.png"),
       plot = grid, width = 20, height = 14, dpi = 300, units = "cm")
ggsave(paste0(outputdir, "/SOCshare.png"),
       plot = SOCshare,    width = 10, height = 7, dpi = 300, units = "cm")
ggsave(paste0(outputdir, "/SOCincrease.png"),
       plot = SOCincrease, width = 10, height = 7, dpi = 300, units = "cm")
ggsave(paste0(outputdir, "/SOCdensity.png"),
       plot = SOCdensity,  width = 10, height = 7, dpi = 300, units = "cm")
ggsave(paste0(outputdir, "/SOCemis.png"),
       plot = SOCemis,     width = 10, height = 7, dpi = 300, units = "cm")
```