### 2022-07-20 ResultNotebooks/results_SOCpointvalidate.Rmd

---
title: "SOC budget analysis in comparsionto point-based measurements"
output:
  html_document:
    df_print: paged
  pdf_document: default
  bookdown::pdf_document2: default
   keep_md: true
---

```{r load}
library(magclass)
library(madrat)
library(mrcommons)

outputdir <- "Output/Images/"

rev         <- "rev30"
scen        <- "histManagement"

SoilCarbon <- readRDS(paste0("Data/", rev, scen, "/SoilCarbon.rds"))
Landuse    <- readRDS(paste0("Data/", rev, scen, "/Landuse.rds"))
```

```{r extract carbon stock for validation year}

year            <- "y2010"
SOCStock_natveg <- dimSums(SoilCarbon[, year, c("actualstate.natveg")], dim = 3)
SOCStock_crop   <- dimSums(SoilCarbon[, year, c("actualstate.crop")],   dim = 3)

SOCcrop_table      <- as.data.frame(setCells(SOCStock_crop, gsub("\\.", "_", getCells(SOCStock_crop))))
SOCcrop_table$Cell <- gsub("_", "\\.", SOCcrop_table$Region)
SOCcrop_table$Native  <- 0
SOCcrop_table         <- SOCcrop_table[-c(2:4)]
names(SOCcrop_table)[names(SOCcrop_table) == "Value"] <- "thisStudy"

SOCnatveg_table      <- as.data.frame(setCells(SOCStock_natveg, gsub("\\.", "_", getCells(SOCStock_natveg))))
SOCnatveg_table$Cell <- gsub("_", "\\.", SOCnatveg_table$Region)
SOCnatveg_table$Native  <- 1
SOCnatveg_table         <- SOCnatveg_table[-c(2:4)]
names(SOCnatveg_table)[names(SOCnatveg_table) == "Value"] <- "thisStudy"
```

### Sanderman validation data (point data)

```{r reformat point validation data for plotting, warning=FALSE}

library(stringr)

  sand  <- readRDS(paste0("Data/", rev, "histManagement/calcSOCPointData.rds"))
  getCells(sand)  <- gsub("\\.", "_", getCells(sand))
  sand_table      <- as.data.frame(sand)
  sand_table      <- reshape2::dcast(sand_table, Region ~ Data1, value.var = "Value")
  sand_table$ID   <- str_extract(sand_table$Region, "[^_]*$")
  sand_table$Cell <- gsub("_", "\\.", str_extract(sand_table$Region, "[^_]*_[^_]*"))
  sand_table      <- sand_table[names(sand_table) %in% c("Cell", "Native", "SOC", "ID")]
  names(sand_table)[names(sand_table) == "SOC"] <- "ValidationData"

  getCells(sand) <- str_extract(getCells(sand), "[^_]*_[^_]*")
  unique         <- unique(getCells(sand))
  names          <- as.vector(outer(c("natveg", "crop"),
                                    c("median", "mean", "max", "min", "q25", "q75"),
                                    paste, sep = "."))
  sand_reduced   <- new.magpie(unique, NULL, names, fill = NA)

  for (u in unique) {
    sand_reduced[u, , "natveg.mean"]   <- mean(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
    sand_reduced[u, , "crop.mean"]     <- mean(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
    sand_reduced[u, , "natveg.median"] <- median(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
    sand_reduced[u, , "crop.median"]   <- median(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
    sand_reduced[u, , "natveg.max"]    <- max(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
    sand_reduced[u, , "crop.max"]      <- max(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
    sand_reduced[u, , "natveg.min"]    <- min(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
    sand_reduced[u, , "crop.min"]      <- min(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
    sand_reduced[u, , "natveg.q25"]    <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1], probs = 0.25)
    sand_reduced[u, , "crop.q25"]      <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0], probs = 0.25)
    sand_reduced[u, , "natveg.q75"]    <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1], probs = 0.75)
    sand_reduced[u, , "crop.q75"]      <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0], probs = 0.75)
  }

  sand_crop_table <- reshape2::dcast(as.data.frame(sand_reduced[, , "crop"]),
                                        Region ~ Data2, value.var = "Value")
  sand_crop_table$Cell   <- gsub("_", "\\.", sand_crop_table$Region)
  sand_crop_table$Native <- 0
  sand_crop_table        <- sand_crop_table[complete.cases(sand_crop_table), ][-1]

  sand_natveg_table <- reshape2::dcast(as.data.frame(sand_reduced[, , "natveg"]),
                                        Region ~ Data2, value.var = "Value")
  sand_natveg_table$Cell   <- gsub("_", "\\.", sand_natveg_table$Region)
  sand_natveg_table$Native <- 1
  sand_natveg_table        <- sand_natveg_table[complete.cases(sand_natveg_table), ][-1]
```

```{r merge validation and output data}

allPoints   <- rbind(merge(sand_table, SOCcrop_table, by = c("Cell", "Native")),
                     merge(sand_table, SOCnatveg_table, by = c("Cell", "Native")))

statsPoints <- rbind(merge(sand_crop_table,   SOCcrop_table,   by = c("Cell", "Native")),
                     merge(sand_natveg_table, SOCnatveg_table, by = c("Cell", "Native")))

correlations <- new.magpie("GLO", NULL, names = unlist(outer(c("median", "mean", "allP"),
                                                            c("crop", "natveg", "allLU"),
                                                            paste, sep = ".")))

correlations[, , "median.allLU"]  <- cor.test(statsPoints$median, statsPoints$thisStudy)$estimate^2
correlations[, , "mean.allLU"]    <- cor.test(statsPoints$mean,   statsPoints$thisStudy)$estimate^2

correlations[, , "median.natveg"] <- cor.test(statsPoints$median[statsPoints$Native == 1], statsPoints$thisStudy[statsPoints$Native == 1])$estimate^2
correlations[, , "mean.natveg"]   <- cor.test(statsPoints$mean[statsPoints$Native == 1], statsPoints$thisStudy[statsPoints$Native == 1])$estimate^2
correlations[, , "median.crop"]   <- cor.test(statsPoints$median[statsPoints$Native == 0], statsPoints$thisStudy[statsPoints$Native == 0])$estimate^2
correlations[, , "mean.crop"]     <- cor.test(statsPoints$mean[statsPoints$Native == 0], statsPoints$thisStudy[statsPoints$Native == 0])$estimate^2

correlations[, , "allP.allLU"]    <- cor.test(allPoints$ValidationData, allPoints$thisStudy)$estimate^2
correlations[, , "allP.natveg"]   <- cor.test(allPoints$ValidationData[allPoints$Native == 0], allPoints$thisStudy[allPoints$Native == 0])$estimate^2
correlations[, , "allP.crop"]     <- cor.test(allPoints$ValidationData[allPoints$Native == 1], allPoints$thisStudy[allPoints$Native == 1])$estimate^2
```

```{r sanderman2, fig.height=8, fig.width=8}

library(ggplot2)
theme_set(theme_bw())
legend_size <- 14
scen_cols   <- c("#785EF0", "#FE6100")
input_cols  <- scen_cols[1:4]

SOCPointValid  <- ggplot() +
                  geom_point(data = statsPoints,
                            aes(alpha = str_wrap("Median (R2=0.13)", 7),
                                x = median, y = thisStudy, colour = factor(Native)),
                            shape = 3, size = 2) +
                  geom_segment(data = statsPoints,
                               aes(alpha = "Min-Max",
                                   x = min, y = thisStudy, xend = max, yend = thisStudy,
                                   colour = factor(Native)),
                               size = 0.5) +
                  geom_point(data = allPoints,
                             aes(alpha = "Point Data",
                                 x = ValidationData, y = thisStudy, colour = factor(Native)),
                             shape = 1, size = 0.5) +
                  labs(x = "Point measurement data (tC/ha) ", y = "Predicted value by this study (tC/ha)") +
                  scale_x_continuous(trans = "log10", limits = c(7, 470)) +
                  scale_y_continuous(trans = "log10", limits = c(7, 470)) +
                  coord_fixed(1) + theme(aspect.ratio = 1) +
                  geom_abline(slope = 1, intercept = 0, color = "grey", size = 0.5, linetype = "dashed") +
                  theme(rect = element_blank(),
                        axis.title.y = element_text(size = legend_size,
                                                   margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
                        axis.title.x = element_text(size = legend_size,
                                                   margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
                        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, size = legend_size),
                        axis.text.y = element_text(size = legend_size),
                        legend.position = c(0.87, 0.2), legend.direction = "vertical",
                        legend.title = element_text(size = legend_size),
                        legend.text  = element_text(size = legend_size),
                        legend.spacing.y = unit(0.1, "cm")) +
                  scale_color_manual(values = scen_cols,
                                     labels = c("Cropland", str_wrap("Natural vegetation", 7)), name = "") +
                  scale_alpha_manual(name = NULL,
                                     values = c(1, 1, 1),
                                     breaks = c(str_wrap("Median (R2=0.13)", 7), "Min-Max", "Point Data"),
                                     guide = guide_legend(override.aes = list(linetype = c(0, 1, 0),
                                                                                 shape = c(3, NA, 1),
                                                                                  size = c(2, 0.5, 0.5),
                                                                                 color = "black")))

SOCPointValid

ggsave(paste0(outputdir, "scatterComparePointData_2010.png"), 
       plot = SOCPointValid, height = 16, width = 16, units = "cm", dpi = 300)
```

```{r sanderman4, fig.height=7.5, fig.width=7.5}

year         <- "y2010"

sand_data    <- readRDS("Data/additional/dataframe.rds")

SOCcrop_data <- as.data.frame(setCells(SOCStock_crop_raw[, year, ],
                                       gsub("\\.", "_", getCells(SOCStock_crop_raw))))

SOCcrop_data$celliso <- gsub("_", "\\.", SOCcrop_data$Region)
SOCcrop_data$native  <- 0
SOCcrop_data <- SOCcrop_data[-c(1:4)]

SOCnatveg_data <- as.data.frame(setCells(SOCStock_natveg_raw[, year, ],
                                       gsub("\\.", "_", getCells(SOCStock_crop_raw))))

SOCnatveg_data$celliso <- gsub("_", "\\.", SOCnatveg_data$Region)
SOCnatveg_data$native  <- 1
SOCnatveg_data <- SOCnatveg_data[-c(1:4)]

test <- rbind(merge(sand_data, SOCcrop_data, by = c("celliso", "native"))[-c(3, 4)],
              merge(sand_data, SOCnatveg_data, by = c("celliso", "native"))[-c(3, 4)])

testcrop <- merge(sand_data, SOCcrop_data, by = c("celliso", "native"))[-c(3, 4)]
testcrop$Value <- as.factor(testcrop$Value)

p <- ggplot(testcrop, aes(x = Value, y = soc)) + geom_boxplot() + coord_flip()

  ### RAW DATA

  points(sand[, , "soc"][sand[, , "native"] == 0], SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0], col = "red")
  lines(c(-1000, 1000), c(-1000, 1000), col = "grey")

  ### MEAN DATA

  plot(sand_reduced[, , "soc_nat"][!is.na(sand_reduced[, , "soc_nat"])],
       SOCStock_natveg_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_nat"])],
       xlab = "Sandermann", ylab = "This Study",
       xlim = c(0, unname(lim)), ylim = c(0, unname(lim)))

  points(sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])],
         SOCStock_crop_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])],
         col = "red")
  lines(c(-1000, 1000), c(-1000, 1000), col = "grey")

  for (u in unique) {

    lines(c(sand_reduced[u, , "soc_crop"] - sand_reduced[u, , "sd_crop"],
            sand_reduced[u, , "soc_crop"] + sand_reduced[u, , "sd_crop"]),
          c(SOCStock_crop_raw[u, year, ],
            SOCStock_crop_raw[u, year, ]), col = "pink")

    lines(c(sand_reduced[u, , "soc_nat"] - sand_reduced[u, , "sd_nat"],
            sand_reduced[u, , "soc_nat"] + sand_reduced[u, , "sd_nat"]),
          c(SOCStock_natveg_raw[u, year, ],
            SOCStock_natveg_raw[u, year, ]), col = "grey")

  }
```

```{r sandermann3}
x <- c(sand[, , "soc"][sand[, , "native"] == 0],
         sand[, , "soc"][sand[, , "native"] == 1])

  y <- c(SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0],
         SOCStock_natveg_raw[getCells(sand), year, ][sand[, , "native"] == 1]) #

  x_clean <- c(sand[, , "soc"][sand[, , "native"] == 0],
               sand[, , "soc"]["CHN.9964", , , invert = TRUE][sand[, , "native"] == 1])

  y_clean <- c(SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0],
               SOCStock_natveg_raw[getCells(sand), year, ]["CHN.9964", , , invert = TRUE][sand[, , "native"] == 1])

  x_crop <- c(sand[, , "soc"][sand[, , "native"] == 0])

  y_crop <- c(SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0])

  x_red  <- c(sand_reduced[, , "soc_nat"][!is.na(sand_reduced[, , "soc_nat"])],
              sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])])

  y_red  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_nat"])],
              SOCStock_crop_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])

  x_red2  <- c(sand_reduced[, , "soc_nat"]["CHN.9964", , , invert = TRUE][!is.na(sand_reduced[, , "soc_nat"])],
              sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])])

  y_red2  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ]["CHN.9964", , , invert = TRUE][!is.na(sand_reduced[, , "soc_nat"])],
              SOCStock_crop_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])

  cor(x, y)^2
  cor.test(x, y)$estimate^2
  cor.test(x_clean, y_clean)$estimate^2
  cor.test(x_crop, y_crop)$estimate^2
  cor.test(x_red, y_red)$estimate^2
  cor.test(log(x_red2), log(y_red2))$estimate^2
```