### 2022-07-20 ResultNotebooks/results_SOCtable.Rmd

---
title: "Table on stock change factors per clinate zone"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrcommons)
library(magpiesets)

outputdir <- "Output/"

rev         <- "rev30"
scen        <- "histManagement"

# untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"),
#       exdir="Data/histManagement", extras="--keep-newer-files ")

SoilCarbon <- readRDS(paste0("Data/", rev, scen, "/SoilCarbon.rds"))
Landuse    <- readRDS(paste0("Data/", rev, scen, "/Landuse.rds"))

```

```{r table scf climatezones}

library(knitr)
library(stats)
library(reshape2)
library(Hmisc)

years             <- seq(1975, 2010, 5)
climateIPCCr      <- readRDS(paste0("Data/", rev, scen, "/IPCC_reduced.rds"))

### calculating stock change factors cropand weighted
SOCDens <- dimSums(SoilCarbon[, years, c("actualstate", "naturalstate")], 
                   dim = 3.3)
SOCDens[, , "naturalstate.crop"] <- SOCDens[, , "actualstate.crop"]

Cshare_cell       <- collapseNames(climateIPCCr * 
                       toolConditionalReplace(
                         SOCDens[, , "crop"][, , "actualstate"] /
                         SOCDens[, , "natveg"][, , "naturalstate"],
                         c("is.na()", "is.nan()", "is.infinite()"), 
                         c(NA, NA, NA)))

CShare_study      <- new.magpie("GLO", years,  
                          outer(getNames(climateIPCCr),
                                c("Q0p05", "Q0p5", "Q0p95"), 
                                paste, sep = "."))

cropMask  <- (Landuse[, years, "crop"] > 0.001)

for (zone in getNames(climateIPCCr)) {

  tempMMask <- setYears(
                 climateIPCCr[, rep(1, length(years)) , zone] == 1,
                 years) & cropMask

  for(y in years) {

    cropTempM <- Landuse[, y , "crop"][tempMMask[, y, ]]
    CShare_study[, y , zone] <- 
    Hmisc::wtd.quantile(Cshare_cell[, y, zone][tempMMask[, y, ]], 
                        cropTempM, 
                        probs = c(0.05, 0.5, 0.95))  
  }
}

getNames(CShare_study) <- gsub("_", " ", getNames(CShare_study))
getCells(CShare_study) <- "GLO"

SCF_table  <- as.data.frame(CShare_study)[-1]

saveRDS(SCF_table, paste0(outputdir, "SFC_climatezone+range.rds"))
```

```{r plot climte map}

climateIPCCr <- readRDS(paste0("Data/", rev, scen, "/IPCC_reduced.rds"))
climateIPCC  <- readRDS(paste0("Data/", re, scen, "/IPCC.rds"))

getNames(climateIPCC) <- c("W-Temp Moist", "W-Temp Dry", "C-Temp Moist", 
                           "C-Temp Dry", "Polar Moist", "Polar Dry", 
                           "Boreal Moist",  "Boreal Dry", "Tropical Montane", 
                           "Tropical Wet", "Tropical Moist", "Tropical Dry")

climateIPCC <- climateIPCC[, , c(5, 7, 3, 1, 6, 8, 4, 2, 9, 11, 10, 12)]

magpie_coord <- toolGetMapping("magpie_coord.rda", type = "cell")
climate2plot <- dimSums(climateIPCC * 
                          new.magpie("GLO", NULL, 
                                     getNames(climateIPCC), 
                                     c(1:12)), 
                        dim = 3)
climate2plot <- merge(as.data.frame(climate2plot), 
                      data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")

library("ggplot2")
theme_set(theme_bw())
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library(RColorBrewer)
library(sp)
library(rworldmap)
library(oce)
library(scales)

legend_size <- 10

map.world <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)
coastline.world <- data.frame(longitude = coastlineWorld[["longitude"]], 
                              latitude = coastlineWorld[["latitude"]])

blucols          <- brewer.pal(9, "Blues")[2:5]
viocols          <- brewer.pal(9, "Purples")[6:9]
redcols          <- brewer.pal(9, "Reds")[6:8]
orgcols          <- brewer.pal(9, "Oranges")[3]
cols <- c(blucols, viocols, redcols, orgcols)

ClimateZones <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(climate2plot, select = c("lon", "lat", "Value")), 
              aes(x = lon, y = lat, fill = factor(Value))) +
    scale_fill_manual(name = "",
                      values = cols, na.value = "white",
                      breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12),
                      labels = getNames(climateIPCC),
                      guide = guide_legend(
                        title.theme = element_text(size = legend_size),
                        direction = "horizontal",
                        title.position = "top",
                        # I shift the labels around, the should be placed
                        # exactly at the right end of each legend key
                        title.hjust = 0.5,
                        label.hjust = 1,
                        nrow = 1,
                        byrow = FALSE,
                        label.position = "bottom",
                        keyheight = 0.5,
                        keywidth  = 2
                      )) +
   scale_color_manual(values = cols) +
    theme(legend.position = "bottom",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          legend.text = element_text(size = legend_size),
          rect = element_blank(),
          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7, 0.7)),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          ) + guides(fill = guide_legend(nrow = 2, byrow = FALSE))
ClimateZones
ggsave(paste0(outputdir, "/Images/climatezones.png"), 
       plot = ClimateZones, width = 20, height = 14, dpi = 300, units = "cm")
```

```{r table ipcc scf}

library(knitr)
library(stats)
library(luplot)
library(reshape2)

climateIPCCr <- readRDS(paste0("Data/", rev, scen, "/IPCC_reduced.rds"))

categories   <- c("IPCC2006.low", "IPCC2006.medium", "IPCC2006.high",
                  "IPCC2019.low", "IPCC2019.medium", "IPCC2019.high" )

CShare_ipcc <- new.magpie("GLO", NULL,  
                          outer(categories, 
                                getNames(climateIPCCr),
                                paste, sep = "."))

y2006m <- "IPCC2006.medium" 
y2006l <- "IPCC2006.low" 
y2006h <- "IPCC2006.high" 
y2019m <- "IPCC2019.medium" 
y2019l <- "IPCC2019.low" 
y2019h <- "IPCC2019.high"

# Medium input regime factor
CShare_ipcc[, , y2006m][, , "tropical_moist"]  <- 0.48
CShare_ipcc[, , y2006m][, , "tropical_dry"]    <- 0.58
CShare_ipcc[, , y2006m][, , "temperate_moist"] <- 0.69
CShare_ipcc[, , y2006m][, , "temperate_dry"]   <- 0.80

CShare_ipcc[, , y2019m][, , "tropical_moist"]   <- 0.83
CShare_ipcc[, , y2019m][, , "tropical_dry"]     <- 0.92
CShare_ipcc[, , y2019m][, , "temperate_moist"]  <- 0.70
CShare_ipcc[, , y2019m][, , "temperate_dry"]    <- 0.77

# Low input regime factor
CShare_ipcc[, , y2006l][, , "tropical_moist"]  <- round(0.48 * 0.92, 2)
CShare_ipcc[, , y2006l][, , "tropical_dry"]    <- round(0.58 * 0.95, 2)
CShare_ipcc[, , y2006l][, , "temperate_moist"] <- round(0.69 * 0.92, 2)
CShare_ipcc[, , y2006l][, , "temperate_dry"]   <- round(0.80 * 0.95, 2)

CShare_ipcc[, , y2019l][, , "tropical_moist"]   <- round(0.83 * 0.92, 2)
CShare_ipcc[, , y2019l][, , "tropical_dry"]     <- round(0.92 * 0.95, 2)
CShare_ipcc[, , y2019l][, , "temperate_moist"]  <- round(0.70 * 0.92, 2)
CShare_ipcc[, , y2019l][, , "temperate_dry"]    <- round(0.77 * 0.95, 2)

# High input regime factor
CShare_ipcc[, , y2006h][, , "tropical_moist"]  <- round(0.48 * 1.11, 2)
CShare_ipcc[, , y2006h][, , "tropical_dry"]    <- round(0.58 * 1.04, 2)
CShare_ipcc[, , y2006h][, , "temperate_moist"] <- round(0.69 * 1.11, 2)
CShare_ipcc[, , y2006h][, , "temperate_dry"]   <- round(0.80 * 1.04, 2)

CShare_ipcc[, , y2019h][, , "tropical_moist"]   <- round(0.83 * 1.11, 2)
CShare_ipcc[, , y2019h][, , "tropical_dry"]     <- round(0.92 * 1.04, 2)
CShare_ipcc[, , y2019h][, , "temperate_moist"]  <- round(0.70 * 1.11, 2)
CShare_ipcc[, , y2019h][, , "temperate_dry"]    <- round(0.77 * 1.04, 2)

getNames(CShare_ipcc) <- gsub("_", " ", getNames(CShare_ipcc))

CShare_study  <- as.magpie(readRDS(
                   paste0(outputdir, "SFC_climatezone+range.rds")))

CShare_median   <- collapseNames(CShare_ipcc[, , y2006m])
CShare_range    <- CShare_study[, "y2010", c("Q0p05", "Q0p5", "Q0p95")] 
CShare_median[] <- CShare_study[, "y2010", "Q0p5"] 
CShare_range    <- round(
                     add_dimension(
                       setYears(
                         CShare_range, 
                         NULL),
                       nm = "This Study (5th - 95th percentile).historical 2010 (5th - 95th percentile)", 
                       add = "model"), 
                     2)
CShare_median   <- round(
                     add_dimension(
                       setYears(
                         CShare_median, 
                         NULL),
                       nm = "This Study (5th - 95th percentile).historical 2010 (median)", 
                       add = "model"), 
                     2)

comparison <- mbind(CShare_ipcc,
                    CShare_median)

comparison_table  <- as.data.frame(collapseNames(comparison))[c(-1, -2, -3)]

range_table       <- as.data.frame(CShare_range)[c(-1, -2, -3)]                          
colnames(comparison_table) <- gsub("\\.", " ", colnames(comparison_table))
colnames(comparison_table)[1:4] <- c("Source", "Management", "Region", "Data")
colnames(range_table)[1:4]      <- c("Source", "Management", "Region", "Data")

saveRDS(list(comparison_table = comparison_table), 
        paste0(outputdir, "SFC_comparison.rds"))
```

```{r kable2image}
library(kableExtra)

getwd()
table <- readRDS(paste0(outputdir, "/SFC_comparison.rds"))

save  <- kbl(table$comparison_table, latex = TRUE, digit = 2) %>%
            kable_styling(latex_options = c("scale_down", "striped"), full_width = FALSE) %>%
              row_spec(7:8, bold = TRUE)

save_kable(save, file = "Output/Images/TableSCF_comparison_new.png")
table
```

```{r table2boxplot, fig.height=3, fig.width=6}

#comparison_table
range_table <- dcast(range_table, Source + Management + Region ~ Data)

cols <- c("#648FFF", "#DC267F", "#FFB000")[3:1]

SCFTable <- 
  ggplot() + 
    geom_linerange(data = range_table,
                   aes(x = Region, ymin = Q0p05, ymax = Q0p95, color = Source),
                   position = position_nudge(x = 0.165), size = 0.6) +
    geom_point(data = comparison_table, 
               aes(y = Data, x = Region, group = Source, color = Source, shape = Management), 
               position = position_dodge(width = 0.5), 
               size = 2) +
    labs(y = "Stock Change Factor",
         x = "Climate Zone") +
    scale_color_manual(values = cols) +
    guides(shape = guide_legend(order = 2),col = guide_legend(order = 1)) +
    theme(rect = element_blank(),
                        axis.title.y = element_text(size = legend_size,
                                                   margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                                   unit = "pt")),
                        axis.title.x = element_text(size = legend_size,
                                                   margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                                   unit = "pt")))

ggsave(paste0(outputdir, "Images/SCF_comparison.png"), 
       plot = SCFTable, height = 8, width = 18, units = "cm", dpi = 300)
```