### 2022-10-21 ResultNotebooks/results_SOCscenario.Rmd

---
title: "Global SOC scenarios"
output:
  md_document:
    variant: markdown_github
  html_document:  
    pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrsoil)
library(magpiesets)

rev         <- "rev30"
scenarios   <- c("histManagement", 
                 "constManagement-1975", 
                 "constManure-1975", 
                 "constResidues-1975", 
                 "constTillage-mixedtill")

 filesNeeded  <- c("SoilCarbon.rds", "Landuse.rds")
# setConfig(outputfolder = "/home/kristine/WORK/Paper/historicalsocbudget/ResultNotebooks/Data/")

#  untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
#        exdir=paste0("Data/",rev,"histManagement"), extras="--keep-newer-files ")
#
# for(item in scenarios){
#     untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
#     exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
#  }

outputdir <- "Output/"
```

```{r Load data and calc globals}

glo_SOCStock <- NULL

for (scen in scenarios) {

  print(scen)
  SoilCarbon <- read.magpie(paste0("Data/", paste0(rev, scen), "/SoilCarbon.rds"))
  Landuse    <- read.magpie(paste0("Data/", paste0(rev, scen), "/Landuse.rds"))
  SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
  SOCStock  <- toolConditionalReplace(
                 dimSums(dimSums(
                   SoilCarbon[, , "state", pmatch = TRUE], dim = 3.3) * 
                     Landuse, dim = 3.2), "is.nan()", 0)
  glo_SOCStock <- mbind(glo_SOCStock, 
                        add_dimension(dimSums(SOCStock, dim = 1),
                                      dim = 3.1, nm = scen, 
                                      add = "scenario"))
}

glo_SOCDebt  <- collapseNames(glo_SOCStock[, , "naturalstate"] - 
                                glo_SOCStock[, , "actualstate"]) / 1000
glo_SOCShare <- collapseNames(
                  toolConditionalReplace(glo_SOCStock[, , "actualstate"] /
                                           glo_SOCStock[, , "naturalstate"], 
                                         "is.na()", 0))

saveRDS(mbind(add_dimension(
                collapseNames(glo_SOCStock[, , "actualstate"]), 
                dim = 3.1, add = "type", nm = "stock"),
              add_dimension(glo_SOCDebt,                    
                dim = 3.1, add = "type", nm = "debt"),
              add_dimension(glo_SOCShare,                   
                dim = 3.1, add = "type", nm = "share")),
        paste0(outputdir, "glo_SOCscenarions.rds"))
```

```{r Carbon inputs (glo)}

CarbonInput  <- read.magpie(paste0("Data/", rev, 
                                   "histManagement/CarbonInput.rds"))
Landuse      <- read.magpie(paste0("Data/", rev, 
                                   "histManagement/Landuse.rds"))
glo_SOCInput <- collapseNames(dimSums(CarbonInput[, , "totalC"] * 
                                        Landuse, dim = c(1))) / collapseNames(dimSums(Landuse, dim = c(1)))

ci2plot <- c("above-ground residues" = "ag.crop",
             "below-ground residues" = "bg.crop",
             "manure from monogastrics" = "livst_chick.crop",
             "manure from monogastrics" = "livst_egg.crop",
             "manure from monogastrics" = "livst_pig.crop",
             "manure from ruminants" = "livst_milk.crop",
             "manure from ruminants" = "livst_rum.crop",
             "natural litterfall wood fraction" = "wood.natveg",
             "natural litterfall leaf fraction" = "leaf.natveg")

glo_SOCInput <- toolAggregate(glo_SOCInput, 
                  rel = data.frame(from = ci2plot, to = names(ci2plot)), 
                  from = "from", to = "to", dim = 3)
```

```{r Plot Management Scenarios}

glo_SOCDebt <- collapseNames(readRDS(
                 paste0(outputdir, "glo_SOCscenarions.rds"))[, , "debt"])

manage_scen <-  c(histManagement   = "histManagement",
                  constTillage     = "constTillage-mixedtill",
                  constManure      = "constManure-1975",
                  constResidues    = "constResidues-1975",
                  constManagement  = "constManagement-1975")

library(ggplot2)
library(scales)
library(gridExtra)
library(cowplot)

theme_set(theme_bw())
legend_size <- 10
scen_cols   <- c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "#FFB000")[5:1]
input_cols  <- scen_cols[1:4]

limits <- c(-42, -34)
breaks <- c(-34, -36, -38, -40, -42)
years  <- seq(1975, 2010, 1)

hSOCemis <- 
  ggplot() + 
  geom_line(data = as.data.frame(-glo_SOCDebt[, years, manage_scen]),  
            aes(x = Year, y = Value, group = Data1, color = Data1)) +
  labs(x = "Year", y = expression(paste("Global \u0394SOC in GtC")), 
       color = "") +
  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
  scale_y_continuous(limits = limits, breaks = breaks) +
  scale_color_manual(values = scen_cols, labels = names(manage_scen)) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size, 
                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.title.x = element_text(size = legend_size, 
                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
                                   size = legend_size),
        legend.position = "bottom", legend.direction = "vertical",
        legend.margin = margin(-10, -10, 0, 0),
        legend.box.margin = margin(-10, -10, 0, 0))

hSOCinput <- 
  ggplot(data = as.data.frame(glo_SOCInput[, years, c(3:4, 2, 1)]),  
         aes(x = Year, y = Value, group = Data1, fill = Data1)) + geom_area() +
  labs(x = "Year", y = "SOC input in tC per ha\n (global average)", fill = "") +
  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
  scale_y_continuous() +
  scale_fill_manual(values = input_cols) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size, 
                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.title.x = element_text(size = legend_size, 
                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
                                   size = legend_size),
        legend.position = "bottom",
        legend.direction = "vertical",
        legend.margin = margin(-10, -10, 0, 0),
        legend.box.margin = margin(-10, -10, 0, 0))

vSOCemis <- 
  ggplot() + geom_line(data = as.data.frame(-glo_SOCDebt[, years, manage_scen]),  
                       aes(x = Year, y = Value, group = Data1, color = Data1)) +
  labs(x = "Year", y = expression(paste("Global \u0394SOC in GtC")),
       color = "") +
  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
  scale_y_continuous(limits = limits,
                     breaks = breaks) +
  scale_color_manual(values = scen_cols, labels = names(manage_scen)) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size, 
                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.title.x = element_text(size = legend_size, 
                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
                                   size = legend_size),
        legend.position = "right", legend.direction = "vertical",
        legend.margin = margin(0, 0, -1, 0),
        legend.box.margin = margin(0, 0, 0, -20))

vSOCinput <- 
  ggplot(data = as.data.frame(glo_SOCInput[, years, c(3:4, 2, 1)]),  
         aes(x = Year, y = Value, group = Data1, fill = Data1)) + geom_area() +
  labs(x = "Year", y = "SOC input in tC per ha\n (global average)", fill = "") +
  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
  scale_y_continuous() +
  scale_fill_manual(values = input_cols) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size, 
                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.title.x = element_text(size = legend_size, 
                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
                                   size = legend_size), 
        legend.direction = "vertical", 
        legend.margin = margin(-140, 0, 0, 0),
        legend.box.margin = margin(0, 0, 0, -20))

hSOCemis
hSOCinput
vSOCemis
vSOCinput

hgrid <- plot_grid(hSOCemis, hSOCinput, ncol = 2, labels = c("(a)", "(b)"),
                  label_fontface = "plain", label_size = 10, align = "hv")

vgrid <- plot_grid(plot_grid(vSOCemis  + theme(legend.position = "none"),
                             vSOCinput + theme(legend.position = "none"), 
                             ncol = 1, labels = c("(a)", "(b)"),
                             label_fontface = "plain", label_size = 10, 
                             align = "v", axis = "tblr"),
                   plot_grid(get_legend(vSOCemis), get_legend(vSOCinput), 
                             ncol = 1, label_fontface = "plain", 
                             label_size = 10, align = "v"),
                   align = "h", rel_widths = c(1.4, 1))

hgrid
vgrid

ggsave(paste0(outputdir, "Images/scenario_horiz.png"), 
       plot = hgrid, width = 15, height = 10, dpi = 300, units = "cm")
ggsave(paste0(outputdir, "Images/scenario_verti.png"), 
       plot = vgrid, width = 13, height = 15, dpi = 300, units = "cm")
```

```{r Plot Initialization Scenarios}

glo_SOCStock_init <- NULL
init_scenarios    <- c("Initial-0_constManagement-1975",
                       "Initial-1901_constManagement-1975")

for (scen in init_scenarios) {
  print(scen)
  SoilCarbon <- read.magpie(paste0("Data/", rev, scen, "/SoilCarbon.rds"))
  Landuse    <- read.magpie(paste0("Data/", rev, scen, "/Landuse.rds"))
  SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
  SOCStock   <- toolConditionalReplace(
                  dimSums(dimSums(
                    SoilCarbon[, , c("naturalstate", 
                                     "actualstate")], dim = 3.3) * 
                      Landuse, dim = 3.2), "is.nan()", 0)

  glo_SOCStock_init <- mbind(glo_SOCStock_init, 
                            add_dimension(dimSums(SOCStock, dim = 1),
                                          dim = 3.1, 
                                          nm = scen, 
                                          add = "scenario"))
}

glo_SOCDebt_init  <- collapseNames(glo_SOCStock_init[, , "naturalstate"] - 
                                glo_SOCStock_init[, , "actualstate"]) / 1000

saveRDS(add_dimension(glo_SOCDebt_init,                    
                      dim = 3.1, add = "type", nm = "debt_init"),
        paste0(outputdir, "glo_SOCscenarions_initConst.rds"))

glo_SOCDebt <-
  mbind(collapseNames(readRDS(
          paste0(outputdir,
                 "glo_SOCscenarions.rds"))[, , "debt"]),
        collapseNames(readRDS(
          paste0(outputdir,
                 "glo_SOCscenarions_init.rds"))[, , "debt_init"]),
        collapseNames(readRDS(
          paste0(outputdir, 
                 "glo_SOCscenarions_initConst.rds"))[, , "debt_init"]))

init_scen   <-  
  c( `histManagement.Initial-natveg`         = "Initial-1901",
    `constManagement.Initial-natveg`         = "Initial-1901_constManagement-1975",
     `histManagement.Initial-eq`             = "Initial-0",
    `constManagement.Initial-eq`             = "Initial-0_constManagement-1975",
     `histManagement.Initial-spinup1510`     = "histManagement",
    `constManagement.Initial-spinup1510`     = "constManagement-1975")

glo_SOCDebt <- glo_SOCDebt[,,init_scen]
getNames(glo_SOCDebt) <- names(init_scen)

library(ggplot2)
library(scales)
library(gridExtra)
library(cowplot)

theme_set(theme_bw())
legend_size <- 12
scen_cols   <- c("#648FFF", "#785EF0", "#DC267F")

input_cols  <- scen_cols[1:4]

limits <- c(-55, -25)
breaks <- c(-30, -40, -50)

startY <- 1975

initSOCemis <- 
  ggplot() + 
  geom_path(data = as.data.frame(-glo_SOCDebt[, seq(startY, 2010, 1), ]),
            aes(x = Year, y = Value, group = interaction(Data1, Data2), 
                linetype = Data1, color = Data2)) +
  labs(x = "Year",
        y = expression(paste("Global \u0394SOC in GtC")),
        color    = "Initialization",
        linetype = "Management") +
  scale_x_discrete(breaks = seq(startY, 2010, 5)) +
  scale_y_continuous(limits = limits, breaks = breaks) +
  scale_color_manual(values = c(scen_cols), 
                     labels = unique(gsub("(.*)\\.(.*)", "\\2", 
                                          names(init_scen)))) +
  scale_linetype_manual(values = c(1, 2), 
                        labels = unique(gsub("(.*)\\.(.*)", "\\1", 
                                             names(init_scen)))) +
  theme(rect = element_blank(),
        axis.title.y = element_text(size = legend_size, 
                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.title.x = element_text(size = legend_size, 
                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
                                                    unit = "pt")),
        axis.text.x  = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
                                   size = legend_size),
        axis.text.y  = element_text(size = legend_size),
        legend.title = element_text(size = legend_size),
        legend.text  = element_text(size = legend_size),
        legend.position = "right", legend.direction = "vertical")

initSOCemis

ggsave(paste0(outputdir, "Images/scenario_init.png"), 
       plot = initSOCemis, width = 15, height = 10, dpi = 300, units = "cm")
```